\PoemTitle{Fever}
\label{ch:fever}
\settowidth{\versewidth}{He drapes a black dinner jacket over his iron shoulders.}
\begin{verse}[\versewidth]
 \qquad  \qquad  \qquad Light breaks

Into circus animals, royalty, confetti.\\*
I'm the woman with the gun.\\*
The woman next door.

My blood is gasoline, heart a chain\\*
Of derailed subway cars and he's on\\*
One of them, laughing.

The nurse's face is oval with concern.\\*
She'll cure me with wool blankets.\\*
I wrestle, wrestle free, walk into the wet

Promiscuous forest, reciting\\*
His name and last words.\\*
He drapes a black dinner jacket over his iron shoulders.

Smiles across the white tablecloth---\\*
There are candles, there are no candles, my pearls\\*
Are cool as rain. I center the crown on his head,

Grab the black champagne bottle\\*
And bash his skull in.  The hospital room\\*
Is empty, the other bed, narrow and dark.

I can't see the ceiling.  My fever swells\\*
Blatant and public, across the entire Northeast.\\*
The impregnable courthouse, the hard-

Ware store, the banks and blistered Woolworth's\\*
All shimmy and wobble, hula down Main Street.\\*
There's no rain, no explanation, 

All the forecasts contradict each other.\\*
Airless night clamps down,\\*
Dawn froths like champagne in the tall windows.

His apartment is in ruins: no walls,\\*
Some porcelain fixtures, the marble floor tiled\\*
Black and white.  Columns encircle

A weed-choked atrium.  I can't see the ceiling.\\*
We wrestle and caress, guzzle the champagne,\\*
Let the candles gutter until we are both

Invisible, lose speech. The knights\\*
Are centaurs, the king's in check, his kingdom\\*
Dotted with armless statuary, black and white.

The forest reeks of decay and must\\*
And heat. The black heat of my black blood\\*
Is suffocating.  A parti-colored chessboard   

Warps across the nurse's face; her voice\\*
Is indistinct, like the hiss and rasp of water\\*
Beneath my bed.  She gives me the moon,

White tablet, and a sweating glass. To help me sleep.\\*
I try to tell her that the moon is empty,\\*
And she's moon-empty too, and too big

To swallow. She strokes my forehead\\*
And leaves, whispering.  The forest is cool now,\\*
The green light stands in columns

Between pillared trees.  His voices echoes quietly\\*
Down in the shaded valley.  I rise \\*
On one elbow, listening.

\hspace{4\vgap} The forest\\*
Roils into jungle: pink monkeys and green,\\*
Glassy snakes tumble from the treetops

Abruptly he is with me, shuddering in my arms\\*
Like a child, his cock inside me\\*
Like another heart, my own, his---

We don't know.  We go back to the beginning.\\*
The young queen wakens in the ruined atrium, her crown\\*
Flaming above her lunar face.

Light breaks.
\end{verse}
