\PoemTitle{Sidereus Nuncius}
\label{ch:siderius_nuncius}
\settowidth{\versewidth}{Where the thunder collides with the mountainside, you ride}
\begin{verse}[\versewidth]
1.\\*
Where the thunder collides with the mountainside, you ride\\*
Up the long, long curve of the highway---\\*
Endless, this afternoon, for you.

2.\\*
What\\*
Freak of fate, what\\*
Fluke or twist? What quirk?

What dream, what\\*
Rage, what skewed\\*
Tarot

Led you\\*
To swerve towards\\*
Death

Unknowing?\\*
\textit{Why?}\\*
The one

Word ghosted\\*
Your death, became\\*
A thousand

Questions, a thousand\\*
Specters of what might\\*
Have been, who

---If only---\\*
Who\\*
You would have been

Had death\\*
Not crowned your\\*
Life just then

Just so.\\*
An accident: was\\*
That moment

Merely a tracery\\*
Sparks and ash\\*
Fallen from the primordial

Flash and roar\\*
That set the\\*
Universe

In motion?

3.\\*
\hspace{4\vgap} Only\\*
The rubric of your blood\\*
Is left as answer, and that

Marginalia proves inscrutable:\\*
It can be made\\*
To say anything.

\hspace{3\vgap} I can search\\*
In vain for apostil or\\*
Colophon: any

Thing to indicate\\*
Origin, intention,\\*
Meaning.

\hspace{3\vgap} I can imagine\\*
The inky washy\\*
Of forest and horizon

Whitened for shivering\\*
Microseconds\\*
By lightning.

\hspace{3\vgap} I can see

The slick asphalt and all\\*
At once lights\\*
In the wrong lane

\hspace{3\vgap} Wrong\\*
Blinding\\*
\hspace{1\vgap} Wrong---

4.\\*
\hspace{3\vgap} But\\*
A wobbly motorcycle driven\\*
Into windshear and rain?

\hspace{4\vgap} I cannot\\*
Think\\*
\textit{Why?}

5.\\*
When I knew you, there were no messages\\*
From the stars, no omens, no fabled heroes\\*
To await or mourn. When I knew

You, you were young. You were almost\\*
A boy, I see now, but your eyes\\*
Were focused on some thing

A boy couldn't know.\\*
In our late afternoon class, the fall\\*
Jimi Hendrix and Janis Joplin died, you read Yeats

A coal-fire sunset flaring in the panes behind us.\\*
In the timbre and rhythm of your voice I heard,\\*
I knew, what you knew, and knew also

For the first time, the elation and severity of Yeats'\\*
Music: the wingbeat of nine and\\*
Fifty swans; the drumbeat

Of the human\\*
Heart at close of day.\\*
There seemed no more to know. Until

Rising from the poem, rising from my own\\*
Resinous heart: the formless\\*
Turbulence, the lover's

Insistent\\*
Gentle\\*
Demands.

6.\\*
Now

As I discover and explore your death, I see you\\*
All gold in the flare of that light, young, not\\*
The bearded Victorian specialist, by the

Evidence of your dust-jacket photo a bit\\*
Stout---\\*
\hspace{2\vgap} Did you forget Yeats? He abandoned\\*
The Victorians and wrote some fierce poems

In middle-age: that's when beauty appeared\\*
Most terrible to him, when he was forced to ask\\*
Its meaning and its use. Your meaning

Has never been so clear or terrible as\\*
Now, when you die\\*
Each time I think of you \qquad then

Shimmer, whole, haloed with autumnal light,\\*
Alive as you never were\\*
Except in memory.

That's the injury, mortal---\\*
But not fatal.\\*
I'll live with the wound.

But how can you not be in this world?
\end{verse}
