\subsection*{Adar Rhiannon (p. \pageref{ch:adar_rhiannon})}
This poem I wrote c. 1980-82. A friend of mine was pregnant, and this is sort of meditation on her
pregnancy, and pregnancy in general: the intimacy of carrying another
being within you, both a part of you and separate, undergoing its own
series of evolutions, from unrecognizable knot of cells to obviously
human.

\subsection*{The Age of Reason (p. \pageref{ch:age_of_reason})}
c. 1989-1990. Based on a memory of going to a gray stone church in West Roxbury when I
was a little girl, probably 7, because I had had it hammered into me
that age 7 was the age at which I was capable of committing a mortal sin
(the age of Reason) and being damned to hell for eternity. The priest's
sermon terrified me, and we came out of the church's gloomy interior
into bright spring sunshine and newly blossomed flowers. Yet I was sure
I would be damned---not for any offense I'd committed, but because
apparently we were all capable of sinning so grievously that God would
torment us for eternity.

\subsection*{Alas Poor Hamlet (p. \pageref{ch:alas_poor_hamlet})}
Composed 2014. An experiment in working in rhyme/slant rhyme. Also, structured as two
interlaced sonnets. I think it's moderately successful---and
fun.

\subsection*{Another Creation Myth (p. \pageref{ch:another_creation_myth})}
Probably early composition, 1977 or so, with later
revisions. Based on one of the early morning bike rides
I used to take, during which my route took me past a field in which
there was a cow, dead on its side, with a calf halfway protruding from
her haunches. She was absolutely still, as was the calf: they must have
died some hours since, and it was just a little past dawn when I came
upon them. The reference to Genesis is figurative---not meant to be
literal.

\subsection*{Appearances (p. \pageref{ch:appearances})}
c. 1988; 08-iii-08 some rev. Not a successful poem. Too
much ambivalence at the heart of it.

\subsection*{Archaeopteryx (p. \pageref{ch:archaeopteryx})}
c. 1982-1984. Simple enough theme: a child's
transformation of reality via her imagination. Gets pretty dark toward
the end. I made wings out of old laundry boxes (not detergent---these
were thin boxes my father's laundered shirts came back in), and I
certainly imagined flying. Didn't make it strictly autobiographical: I
have only a single baby sister here, and my parents virtually never
watched TV together---he worked nights. And in general, he wasn't very
companions. I think my mother must've been very lonely.

\subsection*{As a Child (p. \pageref{ch:as_a_child})}
Probably composed 1988, but my notes indicate it could have been begun
earlier, and rev. til 1988. A fairly straightforward
poem on the horror show that was my childhood. My father's temper ruled
everything, and left me in a state of terror much of every
week.

\subsection*{As Death Approaches (p. \pageref{ch:as_death_approaches})}
1987;~ rev. 07-vii-08. On Cathy Blank's death, from her
parents' perspective. It's an okay poem, not
great.

\subsection*{August Evening (p. \pageref{ch:august_evening})}
c. 1990s. Inspired by a passage in Mrs. Dalloway by
Virginia Woolf: an old woman singing a song that becomes wordless and
primeval. In any case, our civilizations, like our lives, are very
short.

\subsection*{Azrael (p. \pageref{ch:azrael})}
c. 1984; maybe earlier: 1978-82. A meditation on the
fluidity of identity. And how reality is shaped by perception.

\subsection*{Banish Misfortune: A Sonnet, A Jig (p. \pageref{ch:banish_misfortune})}
September 2014. Another fun experiment with rhyming. And
another experiment with the sonnet form. I like
it.

\subsection*{The Blood of Winter (p. \pageref{ch:the_blood_of_winter})}
c. 2002-2003. While my marriage was crashing and
burning. Not subtle, not a poem I think is particularly
accomplished.

\subsection*{Blood Trail (p. \pageref{ch:blood_trail})}
c. 1990-1992. A dramatic monologue.~ Not particularly
accomplished.

\subsection*{The Body of the Dream (p. \pageref{ch:the_body_of_the_dream})}
c. 1978; rev. through 2015-2016. I probably wouldn't
write this poem in anything like its present form today; it edges too
close to something new-agey. But the basic idea---of perception and
reality--- is not that alien to a lot of ideas in modern physics, and
god knows if we live in 10 or 11 dimensions, we aren't seeing things as
they are.

\subsection*{By the Laughless Rock (p. \pageref{ch:by_the_laughless_rock})}
c. 1981-1985. Three poems loosely about mother-daughter
interconnections; about being daughters and mothers; about birth.
Ironically, it turns out my mother was not French-Canadian; her mother
was not French, but Irish. But I can't unweave these poems and ``correct''
their errors---and it doesn't matter: they may not be accurate
according to the literal events of my family history, but the basic
themes survive.

\subsection*{Chaos/Theory (p. \pageref{ch:chaos_theory})}
c. 1998-1999. Written after our visit to Arbroath, in
(north) eastern Scotland, in 1998.


\subsection*{Death in July (p. \pageref{ch:death_in_july})}
c. 1986-1987. One summer, I saw a woman walking along
Rte 1 in Ogunquit or York in Maine, and in my memory, she's dressed all
in white and a man accompanying her holds a white parasol over her. The
sea glittered harshly to their right; I was in a car on their left. She
was not facing me, but rather looked out at the sea. Something in her
reserve reminded me of my mother: my mother wasn't reserved, but death
bestows a kind of reserve, a distance that can't be breached, and I felt
all of this in a flash as I glimpsed her from the car. When someone
dies, for all the power of your memories, you find yourself desiring
their physical presence powerfully.

\subsection*{Deliverance (p. \pageref{ch:deliverance})}
c. 2000. Inspired by the death of one of the professors
at CCC: he died of cancer over Christmas break one year, having let very
few people in on his illness. I remember how shocked I
was.

\subsection*{Desire (p. \pageref{ch:desire})}

\subsection*{Divorce (p. \pageref{ch:divorce})}
c. 1982: rev. often since, including 2016. Based on the dream of a man I saw for several months the year after my mother died.
He turned out to be a nasty piece of work, not worth commemorating with
a name, but his dream was interesting.

\subsection*{The Dogwood Tree (p. \pageref{ch:dogwood_tree})}
c. 1984-85; maybe earlier. This is a poem based in some
of my struggles with the remnants of my Catholic faith: I had regarded
myself as more agnostic than atheist, and around this time I made the
final swerve into full-blown atheism.

\subsection*{Dream House (p. \pageref{ch:dream_house})}
2015. One of the few poems I've written while feeling
more acutely ill. It's not a dream house of the kind you'll find on HGTV: no one lives in such a house, except for me. The blue floor comes from
my memory of my attic room in West Roxbury, with its single dormer
window. That has remained my ideal bedroom ever since, but I'll never
see its like again.

\subsection*{Driving Cross Country in the Summer (p. \pageref{ch:driving_cross_country})}
c. 1993-1994. Written about drive from Yakima to
Corning. A brief glimpse along the way.

\subsection*{Driving Down (p. \pageref{ch:driving_down})}
c. Late 1990s. I was driving back from some therapy
appointment, late for something? Picking up Will or a committee meeting?
It's not a much-worked on poem, and I think what is there is
apparent.

\subsection*{Drunk, Poet (p. \pageref{ch:drunk_poet})}
c. 1989. I'm not of the school that believes being drunk
or high or mentally disturbed improves ones creative work. Drunks are
notorious liars, to themselves, first of all, but everyone else, too.

\subsection*{Easter (p. \pageref{ch:easter})}
c. 1980-1984; rev. 28-iii-16; 3-iii-17. I drew what I
thought was a perfectly reverent picture of Christ on the crucifix and
showed it to Mary, the landlord's daughter: her shock still surprises me
now.

The "wine-dark night" of the opening is one of several vigil nights,
spent sometimes in the hospital, sometimes (as here) at home, waiting
for some turn in my mother's condition. Often I couldn't sleep, and drew
my sisters where they slumped in chairs.

\subsection*{Equinox (p. \pageref{ch:equinox})}
c. 1986? Not much more than a fragment,
really.

\subsection*{Falling Child (p. \pageref{ch:falling_child})}
c. 1988. Nothing profound here. I spent a lot of time in
the park with Alex, and any parent who spends time watching their kid(s)
outside worries about how badly they might injure themselves because of
this or that hazard that the landscape offers.

\subsection*{Fragment (p. \pageref{ch:fragment})}

\subsection*{A Fragment of This Evening's Melody (p. \pageref{ch:a_fragment})}
c. 1977, 1980, and 1982. I've saved this poem primarily
because my mother liked it, not because I think it's a particularly
wonderful poem. It's not something I would write
today.

\subsection*{Friday Morning (p. \pageref{ch:friday_morning})}
c. 1985, shortly before I became pregnant with
Alex. The only love poem I wrote for Bob, who promptly
filed it in a pile of papers on his desk.

\subsection*{HUSBAND AND WIFE FREEZE TO DEATH AWAITING `ALIENS' (p. \pageref{ch:husband_and_wife_freeze})}
c. 1985. I either glimpsed this story or heard it on the
radio, or something very like it.

\subsection*{In Memory of a Beautiful Life (p. \pageref{ch:in_memory_of_a_beautiful})}
May 1980; rev. 3, 4, 5,-iii-2017. As I have remarked
elsewhere, Paul Mariani hated this poem, but I never shared his disdain.

\subsection*{In the Season without Mercy (p. \pageref{ch:in_the_season_without_mercy})}
c. 1990. One of the few autobiographical and narrative
poems I've written, especially in the last few decades; my poems became
more impersonal, for the most part.

\subsection*{In Time (p. \pageref{ch:in_time})}
c. 1989. I think I wrote this after seeing some show on
fertility myths, and so some of that fertility symbolism worked its way
in. I vaguely remember revising it to make it more woman-affirming at
the end---not just one more poem on the loss of a man.

\subsection*{Jungle (p. \pageref{ch:jungle})}
c. 1988. Some get once remarked to me---or I read it in
one of the books I had on the Vietnam War---that it was amazing how
many bullets were in the air during a battle---amazing how you might
not be hit despite this---and someone remarked that there might be a
bullet hole in every leaf.

\subsection*{Litanies (p. \pageref{ch:litanies})}
c. 1975. I still like these, though I wouldn't write
them this way today---yet I don't know how I would write them, so I
haven't really revised them, and there's too much substance to them to
dismiss them.

\subsection*{Lung Cancer, Stage IV (p. \pageref{ch:lung_cancer})}
2015. That rooster lives in the next house down the road
from us, and that he doesn't have laryngitis is a miracle. Just saying.
Also this poem makes reference to another poem of mine, ``\titleref{ch:evening_class}.''

\subsection*{Marilyn Poems (p. \pageref{ch:marilyn_poems})}
c. 1982? 1984? I like these poems; always have. I
vaguely (inaccurately?) recall writing them in workshop and having them
trashed.

\subsection*{Masaccio's \textit{Expulsion} (p. \pageref{ch:masaccio})}
c. 1986-1987. One of my favorite paintings by Masaccio.
He gets anguish.

\subsection*{Medusa (p. \pageref{ch:medusa})}
c. 1982 or earlier; c. 1984-1985; rev. 2016. A poem on
just how scary women's anger is to men, especially if they suspect it
might be justified.

\subsection*{Mythological Bird (p. \pageref{ch:mythological_bird})}
c. 1980-82? Or 1993-95?? I wrote this quite a while
ago. Still one of my favorite poems.

\subsection*{The Nature of Loss (p. \pageref{ch:the_nature_of_loss})}
c. 1988. A poem on my mother's death. Not deep, but worth
saving.

\subsection*{The Oldest Love Poem (Sonnet for Robert Connors) (p. \pageref{ch:the_oldest_love_poem})}

\subsection*{New Lovers \textit{[formerly ``Threat'']} (p. \pageref{ch:new_lovers})}
c. 1988. The poem is better than the guy I wrote it for,
but not so much.

\subsection*{Night Reflections (p. \pageref{ch:night_reflections})}
\subsection*{Noon (p. \pageref{ch:noon})}
c. 1988? Maybe earlier. On a little boy I saw playing
outside the laundromat, which makes me think the date is earlier,
perhaps 1982-1986. The transformation of the world by one's imagination
was my central thought in writing it.

\subsection*{On First Looking into the New Oxford Rhyming Dictionary (p. \pageref{ch:on_first_looking})}
c. 2014-2015. This is a fun ``fun'' poem. I'm fond of it.

\subsection*{Phoenix (p. \pageref{ch:phoenix})}
c. 1978-82; maybe 1984-85; rev. 2016. I wouldn't write
this poem this way now, but I'd have an impossible time revising it more
than I have, and I'm very fond of some aspects, some
lines.

\subsection*{The Poems of Summer and Autumn (p. \pageref{ch:the_poems_of_summer})}
c. 1986-1987. One of the poems that more explicitly
glances at that theme of a mother, a daughter herself, sees a line of
continuity to her daughter: a doubling, yet an
obscuring.

\subsection*{Poem from a Restless Sleep (p. \pageref{ch:poem_from_a_restless_sleep})}

\subsection*{Precipitous Flight (p. \pageref{ch:precipitous_flight})}
c. 1988. Not a bad summation of the reasonableness of
being afraid to fly. But life throws some crises at us that make it
impossible to cave into a perfectly reasonable fear.

\subsection*{Print over the Sink York (p. \pageref{ch:print_over_the_sink_york})}
c. 1998. York in England, not Maine.

Light sometimes obscures. I've always found that ironic, and it
sometimes sets off further meditations on the phenomenon. As here.

\subsection*{Rembrandt Wrestling the Angel (p. \pageref{ch:rembrandt_wrestling})}
c. 1984-85; maybe earlier. Based on a drawing by
Rembrandt.

\subsection*{A Shadowed Poem (p. \pageref{ch:a_shadowed_poem})}
c. 1974-75; revised intermittently since. A poem that
explores the relationships between me, my mother, and her mother:
imagining her life with her mother, her mother's death---but not my
mother's death: this was written 6 or 7 years before my mother died.
Heavily marked by the art history classes I'd taken. Since I didn't take
the Renaissance Art History class until I returned to school, c.
1977-78, then parts of this poem are at least that late. 1974-75 may be
an erroneous guess based on when my mother used to visit me in
Worcester, scenes of which are depicted in the
poem.

\subsection*{Sheer Poetry (p. \pageref{ch:sheer_poetry})}
c. 1990-1992. A series of interconnected dramatic
monologues, each in the voice of one of four people. There are two
couples, and two people---half of each couple---have had an affair.
The affair is over, the ashes have fallen. The woman of the adulterous
couple gets the last word.

\subsection*{Snake Charmer (p. \pageref{ch:snake_charmer})}
c. 1982; rev. 23-iii-16. I don't think this is a great
poem, but I like its rhythms and love the
imagery.

\subsection*{Solitudes (p. \pageref{ch:solitudes})}
c. 1998. I need a lot of time to myself. And like
Cordelia, I want what people to say not to be mere flattering fluff or
bland generalities.

\subsection*{Sunset on Market Street (p. \pageref{ch:sunset_on_market_street})}
c. 1984-85; maybe earlier. The ``Sharon'' of the poem is
fictional. Not a deep poem, but I like it.

\subsection*{Target Practice (p. \pageref{ch:target_practice})}
c. 1982-84: revised frequently after, until final rev. in 2016. It's as good as it can be; it's not a great
poem.

\subsection*{Under Fire  (p. \pageref{ch:under_fire})}
c. 1989-1990. On a vet tormented by war ghosts and
alcoholism.

\subsection*{Unfinished (p. \pageref{ch:unfinished})}
c. 1990s; maybe early 2000s. A story one of my
colleagues at YVCC told me, about the death of the seven-year-old
daughter of one of his students.

\subsection*{Unspoken (p. \pageref{ch:unspoken})}
c. 2002. For an unrequited
crush.

\subsection*{What Birds (p. \pageref{ch:what_birds})}
\subsection*{What I Meant to Ask You (p. \pageref{ch:what_i_meant_to_ask})}