\PoemTitle[Edgar: Death approaches.]{}\label{ch:lear_bu}
\settowidth{\versewidth}{Bearing Cordelia, the sacrifice more fierce in the fulfillment}
\begin{verse}[\versewidth]
\Character{Edgar} Death approaches.

Darkness swallows the forests and rivers.

Darkness settles on the villages and towns.

Darkness spills down the furrows of the plowed fields.

Darkness pools in the byres and sties.

Darkness scales the eyes of the dead.

Darkness inflects the speech of the living.

The years have gathered into darkness\\*
And darkness rises in my heart.\\*
But my limbs are light. I have unfastened

My velvet robes, emptied my purse,\\*
Unburdened and stripped myself\\*
Down to the flesh, if not the soul.

I will cast off the past with the future.\\*
Yet as death and oblivion begin to ferry me\\*
Down their lightless empty rivers

I shake loose, before the final\\*
Dissolution, I shake loose the dusty\\*
Traces of death, like ashes

Crumbling on pink coals, and remark\\*
My skin, delicate with age, my breathing,\\*
Delicate with an urgency I can no

Longer sustain. Sunlight this\\*
Early in spring is still weak,\\*
But warm, and promising heat come

Summer.\\*
\hspace*{3\vgap} The breeze turns cool.  Afternoons still\\*
The shadows grow long, like the teeth of wolves.\\*
His wolfish daughters---murderers, murdered---both

Dead---borne off to lie with their common betrothed,\\*
Dead also: a frame terrible enough for his entrance\\*
Bearing Cordelia, the sacrifice more fierce in the fulfillment

Than in the poetry describing it.\\*
\hspace*{3\vgap} As that flame\\*
Consumed them both, what whisper then\\*
On Cordelia's lips?  What did he hear

When love's silence gathered its full force\\*
And asked nothing of him?\\*
\hspace*{3\vgap} And he,\\*
On the cusp of darkness himself, did he

Have nothing, give nothing\\*
Gazing on her face in love's new light\\*
Winking seraphim flames amid cinders?
\end{verse}
