\PoemTitle[Lear: The butterfly shakes off not the crumpled cocoon]{}
\label{ch:lear_aa}
\settowidth{\versewidth}{Now, one last ceremony, stiff with pomp. Love's oaths.}
\begin{verse}[\versewidth]
\Character{Lear} The butterfly shakes off not the crumpled cocoon\\*
Only, but sheds as well the cramped\\*
Form: it fans out the damp folds in its wings,\\*
Which hamper flight, and flies. \qquad That is the lesson

From Nature, that is the natural progression\\*
Of life.  And so I will shed\\*
Crushing ceremony and the cares of state,\\*
Disencumber myself of bruising crown

Too heavy for my neck, with its pigeon's wobble;\\*
Too heavy for my head, on which even I can see\\*
The blue veins mapped on pink skin.\\*
\hspace*{3\vgap} So I will exchange

Crown and the cares and crosses borne as king\\*
For the remaining years of her kind nursery.\\*
It is fitting that my youngest should usher me\\*
Into the Eden that is Death's first prefiguring

And arabesque. Thus will I sit in the shadow\\*
Of her garden wall, the world shut out,\\*
My eyes closed, the afternoon heat heavy,\\*
But my forehead cool, my limbs at rest,

My thoughts free.\\*
\hspace*{3\vgap} An elderly king\\*
Whose crown has slipped, an old man\\*
Dozing in the shade of a summer after\\*
Noon.

\hspace*{3\vgap} Invisible to all: none will see\\*
My return to the young man I was, conjured\\*
From memory, desiring all that was before me\\*
Before strife and duty and unmastered passion

Delivered me to old age and its mild pleasures.\\*
Now, one last ceremony, stiff with pomp. Love's oaths.\\*
The kingdom divided, peace in the garden,\\*
Death. A natural progression.
\end{verse}
