\PoemTitle[Oswald: Neither my mistress nor her sister]{}\label{ch:lear_bf}
\settowidth{\versewidth}{\hspace*{3\vgap} Mistress and sister---each---}
\begin{verse}[\versewidth]
\Character{Oswald} Neither my mistress nor her sister\\*
Can see it---but I see it.\\*
He worships a goddess but he loves\\*
No one.\\*
\hspace*{3\vgap} Mistress and sister---each---\\*
Flaunts her contempt for all\\*
That stays ambition, and each\\*
Parades her own ambition as summons

To her caress and kiss, her\\*
Promise of submission if not her\\*
Actual capitulation to his will.\\*
\hspace*{3\vgap} But neither\\*
Captures the mercurial devotion\\*
Of her distracted lover, though each simmers\\*
In the lingering intoxication of his\\*
Lust and heat.\\*
\hspace*{3\vgap} Yes indeed, likely war

`Twixt duke and duke---but let them\\*
See to the mutinies within the halls,\\*
The machinations masked by silence\\*
When there should be speech, and speech\\*
Merely ritual in its affection.\\*
\hspace*{3\vgap} Servants\\*
Such as myself are the first\\*
To know, the last to speak

---If we speak at all---\\*
When there is trouble in a household.\\*
We watch. I watched the bastard\\*
Son unseat the father,\\*
\hspace*{3\vgap} then smile\\*
His mirthless smile at either sister\\*
As either suited him; waver, whisper,\\*
Caress without noting which

Whispered in return. He rode off\\*
Without either of them, looking\\*
Straight ahead. I took note.  But\\*
So as to look\\*
\hspace*{3\vgap} disinterested, to look\\*
Like a servant, not someone\\*
Who watches them when they forget\\*
To watch me\\*
\hspace*{3\vgap} I disappear and devise

Some chore or task, some\\*
Service to make me seem\\*
Servile, not necessary but\\*
Convenient: unbidden, I bring\\*
A chalice of wine,\\*
I sharpen the ceremonial\\*
Daggers, I feed the Earl's dogs

When they slobber and beg,\\*
\hspace*{3\vgap} I bide\\*
My time. They forget me\\*
Until they need me. And\\*
---Conveniently---I am there.\\*
The dogs too remember:\\*
They crowd me and whine, sniffing\\*
For meat.  Eying them I think

I'd file my teeth into fangs\\*
If I could.
\end{verse}
