\PoemTitle[Goneril: Let my promises sound like love---let them]{}\label{ch:lear_ad}
\settowidth{\versewidth}{Not wise, and wishes to shrug off care and live}
\begin{verse}[\versewidth]
\Character{Goneril} Let my promises sound like love---let them

Be his world.  Let him give me all\\*
For a story, a child's story.    By my life,\\*
Old fools are babes again.\\*
The world his toy, mapped to his pleasure,\\*
And bloated flatteries, blazoning praise\\*
Of Father and Majesty, accepted with childish\\*
Arrogance.  Oh let him,\\*
Like the king in the fairy tale,

Look at the loom, gaze on empty\\*
Shuttle and shaft and swear he sees\\*
What he does not see but should:\\*
A tapestry of plaited stems and blossoms\\*
Framing lion and burning phoenix,\\*
Basilisk and leopard, leopard\\*
Fastened on cringing antelope,\\*
Antelope dying, dying, dying, ever

Approaching death.    A fabled world\\*
Where death and horror are mere\\*
Embroidery: how beautiful\\*
Death appears, imagined, how elaborate\\*
The arabesques that cramp the living\\*
Into coiled design, frozen pattern:\\*
Gorgeous. But death all the same. So.\\*
Let me tell him I see what he

Does not see but pretends to.  He is old,\\*
Not wise, and wishes to shrug off care and live\\*
As the lilies of the field, to crawl,\\*
He says, toward death unburthened.\\*
But he knows not, nor sees, nor cares to see\\*
What it is to crawl, what monsters\\*
Sniff at his tracks, nor what Death---\\*
And I---will demand, ere we yield

Relief.
\end{verse}
