\PoemTitle[The Fool: On the road past the blackened villages,]{The Fool:}
\label{ch:lear_br}
\settowidth{\versewidth}{Wrath without reason, without justice, without cause.}
\begin{verse}[\versewidth]
On the road past the blackened villages,\\*
The ashes cold under a hard blue sky,\\*
I left, since Death had taken all---the King\\*
I would not call King, and his youngest---\\*
I left, knowing Death was approaching

Winding toward me across the waste\\*
Of muddy fields and empty nights\\*
Intervening.  I left\\*
And rode toward Death.  The war cut down\\*
Earls and dukes, their officers, and soldiers

By the thousands, faceless as dirt, villeins\\*
Dragged into the flames of their lords' petty hells:\\*
Wrath without reason, without justice, without cause.\\*
The world is a furnace, the world is full of death.\\*
So it is. So it is.  Look at it---but look.

Look up.  The tucked and smocked\\*
Underbellies of the clouds---as ornate as any\\*
Courtier's sleeve---catch the saffron and cinnabar\\*
Flame of sunset before charring to ash\\*
With nightfall---\\*
\hspace*{2\vgap} All vanity, all

Transience, to be sure, but\\*
In the instant, so\\*
Stunning everything else\\*
Is forgotten, held\\*
Back at the margin of vision---

Even Death for these few seconds\\*
Still and hushed.\\*
But cold night damp rises\\*
Out of the marsh and a sickle moon\\*
Looks unblinkingly down at me.\\*
That's not the pretty smocking

Of a courtier's sleeve, nor\\*
Mi'lady's ribboned hair\\*
Let down in darkening ripples\\*
As her dress falls and the bed rocks.\\*
That crimson

Is the pan of coals some cruel or\\*
Careless god has thrown down,\\*
And if they light our way\\*
Or sear our feet and immolate\\*
Our children, why

There's no why, no way, to be had:\\*
The path's the same: it's mine:\\*
Join me.  Leave behind those true\\*
Fools: gilded butterflies\\*
Almost solemn

In their vanities:\\*
They fan the perfumed\\*
Chamber air\\*
Unseeing\\*
Unseeing

Though the clouds descending are red as blood.
\end{verse}
