\PoemTitle[At Dawn]{At Dawn\\\textit{(for Rosaleen)}}
\label{ch:atdawn}
\settowidth{\versewidth}{May break her moorings and burn on uncharted waters.}
\begin{verse}[\versewidth]
My mother is no longer herself.\\
Morphine slogs through her veins, dims the light,\\
Dredges up old memories like marionettes.\\
She says she's \textit{ducky-wucky} and spills her soup.\\
Lifting her sagging, bloodless body onto the bedpan,\\
I see that her pubic hair, too, has mostly fallen out\\
And that what is left is colorless and withered\\
Like dead grass. Moved to intensive care,\\
She's left naked: on her legs and half-moon chest\\
Bruises blossom through transparent skin.\\
I imagine that I can almost see her lungs,\\
Two wet orchids, inflating with each glassy breath.\\
A nurse drapes a towel over her.\\
My sisters return with coffee from the cafeteria;\\
We each take our places and begin to wait.

At night, the hospital towers over the black lake.\\
The parking lot glitters with hunched cars.\\
On the other side of the highway, the stars\\
Pinioning the great dog's feet are lost\\
On the neon horizon. Sirius barks after his master.\\
The Pleiades fly in deathless flight\\
Before blue-black Orion, Orion\\
Who once trod our suburban backyard,\\
Glinting in his starry belt and lion's skin.\\
Framed in the window by her narrow bed, the fires\\
Of the ice-fishers on Lake Quinsigamond\\
Dance like sparks against the mottled dark.\\
My father is trying to reason with the doctors.\\
But she is lost deep in her coma and tonight\\
May break her moorings and burn on uncharted waters.

The scent of eucalyptus is heavy by her bed:\\
A salve or disinfectant that dispels\\
The vegetal musk of her body.\\
The digital faces of the life-support monitors\\
Flicker toward silence. What use\\
Are these instruments? Light\\
Floods toward the center of the city:\\
Two long arms of traffic course\\
In from the dark perimeter.\\
There's a bridge over the narrow lake,\\
And silence under the bridge and darkness.\\
My father ceases reasoning and joins us.\\
The priest says \textit{Bless your daughter,}\\
\textit{Rosaleen}. There's no more blood.\\
She looks out the window into the night.

Is there a luminous Christ balanced\\
On the stone sill? Her mother's cameo profile,\\
Inclining over her daughter's mirror face?\\
Blackness?\\
Does she recite the litany of chores never\\
Quite finished, and push them, perpetually,\\
Towards a completion ever more final?\\
She folds the white napkins into perfect\\
Triangles. There. There. There. That's nice.\\
The universe is Euclidean, and painless.\\
The tube entering her mouth drags her face\\
Down, shunts her tongue awkwardly up.\\
Jack-o-lantern: she once made this face\\
To make us laugh. We weren't afraid. The darkness\\
Was haloed with street lamps, the ghosts

Were sheeted children. Blackness?\\
The inhuman rasp of the respirator\\
Measures the night. The ward is quiet;\\
Nurses make their murmuring rounds.\\
This one with the round face and black hair\\
Is my favorite. She has children.\\
She will go home to them at dawn. When\\
I turn from my mother at 4 or 5 a.m., yielding\\
Her bloated hand to my brother,\\
She no longer looks frightened. I see\\
Her reflection in the black window: the fires\\
Of the ice-fishers crown her head and heart.\\
Slowly, scorpion dawn emblazons\\
The ragged horizon. Slumped about her bed,\\
Sleeping, her grey children are washed

In rose light. The clouds above the lake\\*
Are flame and angels' hair: the ice-bound lake\\*
Burns. We have come to a dawn\\*
Beyond which we can see\\*
No other.
\end{verse}
