\PoemTitle{Death in Life}
\label{ch:deathinlife}
\settowidth{\versewidth}{To enter the darkness that astounded, night after night,}
\begin{verse}[\versewidth]
You wanted to be somewhere else.\\*
Wherever you were, you desired\\*
Not to leave, but to be\\*
Past, over, everything---to wander\\*
The uncolonized territories, beyond \\*
The orange hills, unframed by windows---\\*
To enter the darkness that astounded, night after night,\\*
Ascending from the grid of streets and lights\\*
Delimiting your suburban life.\\*
You told me stories about your childhood\\*
But stopped before you reached the end.\\*
You recited fables without morals.  Your youth

Seemed spare and tragic, yet not sad.\\*
Not merely sad.  But somewhere, somehow---\\*
A sundering: that fabled life went dark.\\*
Years of silence---then out of that gulf slid\\*
Children, not knowing, not seeing\\*
The blackness they evolved from: one by one,\\*
Reflecting your face in fragments,\\*
We arrived, bloody but whole,\\*
With no past, no history.  The gift\\*
Of oblivion, offered to you in our hands.\\*
And so you talked of yourself\\*
As if you had ceased to exist.	You died

Before you died.   When death arrived, it was a kind\\*
Of child, curled within you\\*
And grasping.   Each day more\\*
Fragile and translucent, your body slowly unveiled\\*
That inner flame, sacrificial, obliterating---\\*
Consuming away future and past, \\*
Illumining your flesh but darkening\\*
Your spirit. When you reached\\*
The end, your children surrounded you,\\*
Awaiting a story, any story,\\*
No matter how brief, if only we might\\*
Hear our names in their original innocence---

Before we read the darkness that haloed you\\*
Or felt that shadow flickering in our bones.
\end{verse}
