\PoemTitle{Requiescat}
\label{ch:requiescat}
\settowidth{\versewidth}{Shopping with my mother: she'd like to buy tight}
\begin{verse}[\versewidth]
Gold Roman numerals flare\\
Then darken: the late afternoon light\\
Slants over the spires, the carved red\\
\textit{Fleur-de-lis} of the clocktower.\\
It seems later than it is. It is February.\\
I think: March. \textit{Not this March.}

Not this spring, I walk downtown\\
Shopping with my mother: she'd like to buy tight\\
New shoes for Easter.\\
In the display windows,\\
Black, silver, and red high-heels\\
---For New Year's celebrations---\\
Are replaced by wafer-thin sandals, pink\\
And grey and bone.

At the florist, we linger, then take\\
Tulips and ferns stapled in tissue, emerge\\
On a breath of eucalyptus and carnation.\\
Water runs out from the dirty snow banks\\
Piled ragged along Main Street. \qquad We cross,\\
Select a wine, a baguette,\\
For dinner. \qquad Passing\\
The \textit{Youth Shop}, I shrug

At the child-sized bunnies, hoarding\\
Their cache of pink and canary eggs,\\
Waving and wagging from their nest of green\\
Cellophane grass: hoarding and wagging, wagging\\
And waving: \textit{How trite, how false.}\\
My mother shrugs\\
At me, reminds me the window\\
Is for children: when I was a child---

The bells ring: it's five o'clock\\*
And dark. Snow begins as rain\\*
But slowly changes over.\\*
I walk home from the library, dreaming\\*
Of bare legs, cotton blouses, my nipples\\*
Hardening beneath my husband's hand. Mother,\\*
When I was a child, I thought the Resurrection\\*
Meant sugar spun into images,

Dresses too thin for pagan spring.
\end{verse}
