\PoemTitle{A Mother Watching her Daughter}
\label{ch:amotherwatching}
\settowidth{\versewidth}{There is no green light, no green world, no daughter}
\begin{verse}[\versewidth]
I

 ``The phrase 'the green world' has echoed\\
In my mind so many times this spring---\\
I can no longer think\\
Where it originated.  Shakespeare?   I imagine \\
Shakespeare's world a green world\\
Bordering a dark heath where Lear raged\\
And all his daughters childless died.''

II

A ridge of lighted heath---My mind is \\
Like a ridge of lighted heath--- alive, glancing,\\
Devouring. And one dark figure\\
---A fool, a king---dances along the edge\\
---A mad king, a father\\
Silhouetted on the burning horizon.\\
There is no green light, no green world, no daughter\\
Where he dances enraged.''

``Yet there is a green world---''

III

``---My daughter howls. Sobs and howls.\\
She has begun to inhabit my poems\\
Even as every day she more profoundly inhabits\\
Herself. And in the process\\
---Her mind leafing, blossoming---\\
She separates from me even as I incline\\
Over her crib where she howls\\
And press her against my breast.''











IV\\
``A burning bush which  is not consumed---\\
Leafing, blossoming, spiraling\\
Into green flame---this\\
Is the daughter engendered\\
Last summer and born\\
In the spring.   She is a fool\\
Such as teased Lear.   Her nonsense sheds light.\\
Her father's face grows simple under her glance.''

V\\
``Her fascination with the ordinary world\\
Fascinates me: she teases me\\
With questions: What do you see?\\
Who are you? Where from?\\
And where are you taking me? O brave\\
New daughter, oh brave new world

That has such wonder in her.''
\end{verse}
