\PoemTitle*{Christmas Light}
\settowidth{\versewidth}{Strung in the maple saplings along Main Street. Like}
\begin{verse}[\versewidth]
In December, there are Christmas lights\\
Strung in the maple saplings along Main Street. Like\\
Unravelling lace they embroider the edge\\
Of the night’s mantle. In the windows\\
Of houses and apartment complexes, electric

Candles glow steadily. Blue, or orange.\\
Shoppers jam the downtown, prodigal \\
Sons and daughters come home.   Festivity crowds\\
Every turn; indoors or out,\\
One is never alone.

			But January \\
Is truly dark.   All the decorations\\
Are stripped, the sun is low, the long shadows\\
Of people and animals devour the weak\\
Afternoon light.   On the Feast

Of the Epiphany my mother is admitted\\
For the last time to the hospital\\
By the frozen lake.   Bleeding,\\
Unable to speak, quickly sinking\\
Into her body she is rushed

Into Intensive Care.\\
			The air is so cold\\
In January, so cold, so bitterly cold\\
That it seems impenetrable, inhuman, planetary—\\
A vast glacier, a splinter\\
In each of our hearts.

			The day after\\
Saint Valentine’s she dies still\\
Uncomprehending, green eyes still\\
Open.   A full month\\
Before the equinox, the day of her funeral,

\\
[new stanza]



The air is moist and balmy, the earth\\
Green in patches, spongy, fragrant.\\
At the edge of the cemetery is a leafless\\
Black elm, filled with starlings.\\
Their discordant chittering sounds

Like farm machinery, clanging, whirring, reaping\\
While we stand \\
About her grave, in mourning, waiting.
\end{verse}
