\PoemTitle{No Time, No Place}
\settowidth{\versewidth}{To the black hills coiled by the roadside, the}
\begin{verse}[\versewidth]
The highway uncoils before me\\
Like a snake, a black

Snake\\
Wet, glistening

With rain.     In receding\\
Rows along each side, tall

Mercury-vapor lamps, their\\
Lights haloed, unfocused, as

If not seen\\
But remembered.     Beyond

Is a lush and endless\\
Blackness

In which earth and sky\\
Fall away.     The bus

Glides down\\
The hill, around

The blind curve, hissing\\
A cloud of silver droplets, plummeting

Into the stone-dark night.     I fall\\
Asleep, soothed

By the rubbery slap-\\
Slap of the windshield wipers.

The windshield streams with tears.\\
I have

Undertaken a pilgrimage to the hospital\\
Made of black stone:\\

I walk, I walk, I walk\\
Down fluorescent corridors

Into the heart:\\
The door to her room is white,

Opens on soundless\\
Hinges.     Cancer has made her

Translucent: she levitates\\
Above the bed,

Swaddled\\
In sheets,

Luminous\\
A saint---

Or the vision of a saint.\\
In the northern quadrant of her window

The lights on the highway\\
Burn.     Rain-black

Slithers down the glass.     My breath\\
Steams the window, the engine

Hums, the bus\\
Cleaves the night in two.

The highway\\
In a tightrope

Strung\\
Above the amniotic landscape.

We enter the night as if we own it:\\
We have entered the bloodstream of a giant.

Behind us, our taillights glow like the embers\\
Of a ritual fire, an offering

To the black hills coiled by the roadside, the\\
Black fields, the bottomless sky.

We follow\\
The white slashes as they leap

---At us---\\
---Back---

Into the night.
\end{verse}
