\PoemTitle{Milano / Pinocoteca Brera}
\label{ch:milano_pinocoteca_brera}
\settowidth{\versewidth}{The window, of course, is narrow, \& the view}
\begin{verse}[\versewidth]
GESU CRISTO ALLA COLONNA\\
Bramante

Over his right shoulder\\
A small window enclosing\\
A delicate landscape, tinted blue\\
And olive-green.\qquad There is a lake\\
Or inlet, a pier, boats, silk tents;\\
The far shore rises in a gentle slope\\
Bordered by clumped trees on two sides;\\
Beyond, mountains pale \& sheer\\
As glaciers dissolve\\
Into the ice-blue dawn.

What landscape is it?  You tell me\\
The scenery of southern Italy\\
Really looks like this: steep \& surreal\\
Peaks, little patchwork fields\\
Bordered by olive and cypress,\\
Geometric ruins embroidered\\
By wild herbs \& flowers,\\
Jewel-like insects with glassine wings.\\
Then tell me: who lives\\
In those silken tents, perfectly mirrored\\
In the lake's still waters?  Who steers\\
The narrow boats?  Who gathers\\
The olives?  Watches from the piers?

Is the earth really so still, so empty?\\
The window, of course, is narrow, \& the view\\
Distant. If there are women harvesting\\
On that far slope, or men navigating\\
The ethereal lake, they are obscured\\
By the miniature scale,\\
And by the distance between them\\
And us.   The light falling across the lake\\
---Or inlet---cannot reveal them.\\
But it illumines the body of Christ, muscular\\
And youthful, though his face is aged.

This Christ is clearly a man---\\
His arms, shoulders, chest\\
Those of a carpenter---\\
His face is divided in half\\
By ragged chiaroscuro.\\
A rope is knotted around his neck\\
Another cuts into the flesh\\
Of his left arm: these bind him\\
Bodily to the engraved column\\
But his troubled gaze follows\\
Something to our right.\\
Is there a cry on his parted lips?\\
What cry?
\end{verse}
