\PoemTitle{Milano / La Scala}
\label{ch:milano_la_scala}
\settowidth{\versewidth}{The maestro ascends, bows, pauses. We attend.  Violins,}
\begin{verse}[\versewidth]
\epigraph{\textit{Balletto su musiche di Gustav Mahler: ``Das Klagende Lied'' e ``Adagio'' dalla Sinfonia n. 10}}{}

LA SCALA

Is all scarlet, white, \& gilt. Velvet\\*
Tiers ascend to the dusky ceiling;\\*
Mirrors in carved frames line the lobbies, reflect\\*
Solemn patrons gliding across gold carpets.

The \textit{secunda galleria} is full\\*
---Mostly--- of students, a few soldiers from Como.\\*
The maestro ascends, bows, pauses. We attend.  Violins,\\*
Cellos---the strings spiral up the melody, the slow

Strain intensifies.  Mahler's unfinished tenth,\\*
Sheared off by his death, is here compressed\\*
Into one movement, gathering\\*
Toward a chastened sadness in the final bars

Of the ``Adagio.'' We are left in darkness.\\*
There is a pause, then intermission:\\*
The hush shatters, the audience splinters,\\*
Smithereens of color eddy across

The theater floor.  The lobbies swarm.\\*
Golden cornets plead for the return\\*
To order; the house lights flash \& sink.\\*
Below, the soprano, mezzo, \& tenor

Line up in front of the obscured stage.\\*
A ballet emerges from the dark air.\\*
Two dancers, two men, take the stage.\\*
Wreathed in shimmery, scaly green, one

Slender as a snake; the other\\*
Big, severe, anguished,\\*
Robed in red \& grey.   They circle\\*
Each other.  Gathering like a dark reef

Behind the corps de ballet, rising, sinking\\*
In a breath, as a single breath: the air flushes aqua,\\*
Green, ultra-marine blue. Golden courtiers\\*
Become black silhouettes, the Queen-to-be

Invisible, swathed\\*
In Stygian night.\\*
The men mime their anger\\*
While music closes on them

Like a fist.
\end{verse}
