\PoemTitle{Night Passage Through Broken Sonnets}
\label{ch:night_passage}
\settowidth{\versewidth}{We're ascending the Alps, sneaking through}
\begin{verse}[\versewidth]
 \hspace*{2\vgap} The train\\
Burrows into the night, the obscured

Continent. \qquad I press my face\\
To the glass, \qquad trying to outline

The lineaments, the silvered\\
Profile \& runic anatomy of this

New\\
Geography

But the night\\
Deepens, a tunnel

Without sides.\qquad Toward\\
Midnight I doze,

Then awaken when the air grows chill,\\
Glints with splinters of snow

Floating in silence.

We're ascending the Alps, sneaking through\\
Switzerland at 2 or 3 a.m.\qquad A little

Moonlight reveals a marble\\
Landscape---contours softened

By snow, three feet deep\\
On the roof of the cottage, billowing

Round the low, lightless train station.\\
Glittering hollows, sheer slopes

Follow. The constellations\\
Icy blue in the Stygian heavens,

Are ciphers, all\\
Except Orion

Dragging his heels\\
On the western horizon. \qquad Lyra

---The mid-summer avatar\\
Of lovers lost \& sought, lost

And seeking---has slipped\\
Down its descending

Arc, below the toothed and\\
Spired edge of mid-

January's night\\
Sky. The train, as if

Pressed by the weight\\
Of winter's dark, descends

Black valleys into Italy.\\
In my compartment, there are no

Other passengers---just slumped\\
Twists of tossed rag dolls, heads

Lolling, shadowy

Faces bland \& sagging---\\
Murmurs through the wall

Sink back into silence. \qquad If\\
I sleep, I do not think

I do, but watch\\
The world

Rocking in my window\\
As it slowly focuses:

Edges, then forms,\\
Appear. Diffuse pink light

Filters through fog,\\
Backlighting a pony that grazes\\

Near a leafless red bush.\\
The fog lifts---a meter---two

And snatches of scenery\\
---Stucco houses,

Vineyards---\\
Leap forward

Are swept\\
Past. \qquad Dawn

Opens over the blond\\
Fields, silhouettes

The twin rows of\\
Cypress trees

Kindles each\\
Lucent blade

On the new\\
Graves.
\end{verse}
