\PoemTitle{The Golden Mile}
\label{ch:the_golden_mile}
\settowidth{\versewidth}{Off the Golden Mile: behind us, night closes over}
\begin{verse}[\versewidth]
Avenue of jittery neon, lights\\
Throng up \& down\\
The Golden Mile: Route 9, thirty miles\\
West of Boston. The preposterous chandelier\\
Visible in the Chateau de Ville\\
Gaudy as a Vegas showgirl.\\
But the bus stop at the edge of Shopper's World\\
Is dark: the mall itself dark: a few ghostly\\
Mannequins stalk the windows, are held\\
At bay by glass. You\\
Pace near me,\\
Singing softly, softly\\
Catching my free hand.\\
Is anyone home?  It's late.  The phone rings.

My sister answers, sleepy.\\
You're singing, still. She hears you:\\
\textit{Who is it?} My name\\
Is enough---she will come.\\
Until then I try to explain\\
My family, then describe the route\\
To my parents' house, to the bedroom I frescoed\\
When I was sixteen: there are doves\\
And peacocks, psychedelic suns \& moons,\\
Constellations of images without narrative\\
On every wall. I ask if you know the words\\
To a song I cannot remember.\\
I ask to distract you, I ask

To forget, not remember.\\
My sister arrives.  We all cram\\
Into the front seat. She asks you\\
About your yellow \& azure country,\\
About Venice, its thousand-year-old\\
Stone lions, salt-eaten palaces, one-hundred\\
And seventy-four canals, thick,\\
She imagines, with black gondolas, the songs\\
Of gondoliers. Were you singing\\
\textit{A bacarole? Venice is like that}\\
You reply---\textit{\& not like that}---\\
\textit{It's just as you imagine}\\
\textit{And it is}---you shrug---\textit{my English}---

\textit{I}---your voice trails away.\\
My sister nods vaguely, \& turns\\
Off the Golden Mile: behind us, night closes over\\
The pink \& gold shimmer, the white marquees.\\
Suburbia is asleep: the clumped trees,\\
Box-like colonials \& ranch houses,\\
Oval shrubs in endless symmetry\\
By door after door---all\\
Are black. We approach silently. A silver key\\
Unlocks the door.  We are home.\\
But the sky, moonless and inky---here---\\
And now---now over Venice flushes\\
Rose to straw-color. There is another world.
\end{verse}
