\PoemTitle{A Smooth Stone}
\label{ch:a_smooth_stone}
\settowidth{\versewidth}{Beach at sunset.   I have been combing the lower beach}
\begin{verse}[\versewidth]
1.\\*
Even as I walk, I begin to walk\\*
In memory---of the clouds, of the rose-colored\\*
Beach at sunset.   I have been combing the lower beach

For a smooth white stone,\\*
An oval to nestle in my palm \& there\\*
Grow warm.           A simple object,

But not to be found.

2.\\*
Two women are approaching: they walk\\*
Hand in hand.  One infrequently stoops\\*
To collect the porcelain halves of clamshells

Picked clean by gulls and waves.  Her gaze\\*
Follows the fair-weather clouds as they drift and\\*
Billow over the dunes: they gather

Above the shore's blond lip but do not spill\\*
Seaward.                                          Under\\*
The caress of the off-shore breeze, her skin

Shrivels into gooseflesh.    The ocean\\*
Appears clear sapphire; the blown sand\\*
White.     But in her dreams,

The sharks in the belly of the ocean\\*
Are unsleeping---\\*
Sleepless, circle toward the pinnacled surface.

\hspace{4\vgap} Walking erect\\*
The other woman carries herself like a dancer:\\*
Spine a rope of seaweed, rock-\\*
Anchored, fluid as the brine that bathes it.

About her ankles\\*
The salt water eddies---\\*
The serrated ocean gnaws

At the gristle that binds her, muscle\\*
And bone.   Her heart in her breast is a shark, hooked\\*
On the barbs of her hot blood, sleepless.

3.\\*
In the rose-colored light\\*
In the long hour just before \\*
Sundown, strands of

Children, couples, a few\\*
Lone figures stroll the inter-\\*
Tidal flats; some

Accumulate at the \\*
Ocean's foaming edge, washing\\*
To and fro with

Each wave    even   as the\\*
Submerged beds of pungent seaweed\\*
Wash to and fro.     To

My right, on the sand, the cast\\*
Shadow of a white shell shifts\\*
And lengthens in counter-

Point to the sun; the leaping\\*
Waves rattle in the ocean's\\*
Narrow throat.     \\*
\hspace{4\vgap} While not looking

I find a green, translucent\\*
Grasshopper, sodden, glued\\*
To the sand by its own 

Weight. Twitching, half-alive, it\\*
Has been pierced by the spears of\\*
Evening light, is lit

As though lit from within: carved glass\\*
Lantern, it flickers, mis-\\*
Placed on the\\*

\hspace{3\vgap} Unraveling\\*
Tidal\\*
 \qquad Sands.
\end{verse}
