\PoemTitle{Evening Class}
\label{ch:evening_class}
\settowidth{\versewidth}{On the far side of the bridge, approaching Northampton,}
\begin{verse}[\versewidth]
1.\\*
On the fifth level of Herter Hall, twelve\\*
Students are gathered at a table too long\\*
---Almost---for the room.  Late fall,\\*
Early evening---the classrooms \\*
Deserted surrounding this one.  \qquad Out

The window I can see headlights\\*
Swoop down the entrance ramp to 116\\*
A mile distant.   Slumped\\*
Liked the slabs and lintels of ceremonial\\*
Ruins, shadows on the sloping olive landscape.

With a rising, then a falling \\*
Voice, the last lines of a poem are recited:\\*
How did you mean this to be read?\\*
I can hardly make out the highway,\\*
Broken like Morse code

Across the dusky hills.   Did you intend\\*
To be ambiguous?   Yes.  No.\\*
I didn't realize you would look at it\\*
Like that. \qquad The instructor \\*
Begins to gather his papers

---It is almost time---

2.\\*
If I wandered out of the woods and stood\\*
---After a thousand years, 

Tonight---\\*
Hemmed between cryptic hills

And highway roped with light\\*
I would see

The southwest dormitory\\*
Complex: rectangles of light 

Pitched skyward \\*
Helter-skelter 

In rapid-\\*
Fire 

Counterpoint\\*
To the black

Forest clumped at its feet.\\*
Magical construct, requiring a god

To explain---     to inhabit--- \\*
To inherit it.

3.\\*
On the far side of the bridge, approaching Northampton,\\*
The bus eases to a standstill    then proceeds\\*
Forward at a crawl.\\*
Why have we stopped?   What happened?   Was someone\\*
Killed?

Ahead of the bus, a long stream of red\\*
Tail-lights; far ahead, two blue lights\\*
Oscillating like blips on a radar screen.\\*
Off to the side, an ambulance---the \\*
Word, ``AMBULANCE,'' spelt

 Backwards beneath its windshield.\\*
How do you mean this \\*
To be read?     One car\\*
Braked at a violent angle across the right lane\\*
The other

Crumpled like tinfoil, all the windows\\*
Splintered, glass sparkling in the shadows\\*
Where cornfields border the highway.\\*
Oncoming cars fill the open lane: \\*
A mackerel pattern ripples 

Through the bus's interior, across\\*
Our animal faces.     Flushed\\*
From the cornstalks: a cat---\\*
Eyes reflecting small, perfect purgatories\\*
Back at the headlights.  Two people

In flannel shirts and blue jeans\\*
Emerge from the cornfields and stand\\*
By the side of the road, hands\\*
In their pockets, spectators---\\*
Or---How do I know?---genii 

Of the cornfields, small gods, the\\*
Souls of the dead, surprised to see us.
\end{verse}
