\PoemTitle{Fossil Life}
\label{ch:fossil_life}
\settowidth{\versewidth}{Of their spiky relatives, it is odd how I ceaselessly eye,}
\begin{verse}[\versewidth]
Consider my irrational dread of spiders:\\*
All legs \& eyes (as many as eight, each\\*
Watching me with incomprehensible \\*
Arachnid thought). And consider the panicked \\*
Splash-dance horseshoe crabs inspired:\\*
Wholly seized by fear when they materialized\\*
At my toes, in the shallows of Green Harbor:

I was nine or ten. They were ancient, \\*
Ratchet-legged, \& moved blindly\\*
Towards me from some world preceding Eden, a world\\*
Before human gods or forms, a world\\*
Indifferent whether I lived or died.\\*
In those unblinking eyes an untranslatable\\*
World looked out.  But did it observe

Me?\\*
 \qquad Seeing how I shrink from the gaze\\*
Of their spiky relatives, it is odd how I ceaselessly eye,\\*
Ceaselessly finger, these trilobites, even\\*
In photographs, their millepedal chorus lines\\*
Of pleura, legs, \& spines, the bare, blunt\\*
Head, thick thorax, tail, all armored\\*
And shielded, fierce as a samurai, all

In vain.  For here sits one, millennia from home,\\*
In the shallow sunlight on my oak shelves.\\*
The local biota are indifferent.\\*
Silvery CDs dispersed in rainbow-streaked\\*
Schools clutter the desk.  The hard drive\\*
Hums \& clicks its occasional\\*
Commentary. Blue-green shadows





[new stanza]



Fill the window, pool on the floor\\*
Where forgotten texts form low shoals. But the leaves\\*
Of this morning's book lift \& fall in currents of light\\*
And thought. It's not a world the trilobite could read\\*
Were her reef-knot brain sparked to activity\\*
So many tens of hundreds of millions of years\\*
After it flashed into darkness \& her blind armor

Laid in the silty ocean floor as in a crypt.  Still\\*
To my sleepy eyes, in slow motion this one\\*
Specimen oars her zig-zag way in dragon-fly\\*
Light without a thought I might recognize\\*
As thought. A dream swaying on its tidal stem:\\*
What strange tropic ocean\\*
---All gold all blue all salt all motion---

Did she rise through to arrive\\*
Here, in my airless study?  And what still\\*
Unevolved, uninvented language will read me\\*
---Millennia more than I can imagine---\\*
As hieroglyph, as dancer---choreography \\*
Lost, music lost, choreographer\\*
Lost, though she appears frozen on her stage

Eternally motionless, eternally listening?
\end{verse}
