\PoemTitle{Leaving Butte, Montana}
\label{ch:leaving_butte}
\settowidth{\versewidth}{Dreams of its copper kings, its Mycenaean past, the past}
\begin{verse}[\versewidth]
It's necessary to leave when the streets \\*
Are blue \& people's faces harsh\\*
From lack of sleep. In this ashen light\\*
No one's graceful.  But we each recognize\\*
The other---we're the earnest, the true\\*
Travelers, the ones with a purpose,\\*
The ones that risk the most awful distances.

At this ungodly hour, the restaurant\\*
Is barred, marble veneers wiped clean, grills cold.\\*
The one bulb dully fluoresces. Derelict \\*
Office buildings loom over narrow avenues\\*
And night lingers in those canyons\\*
Even as, miles to the south \& east, gold\\*
Washes the lodgepole pines where they crowd\\*
The steeps of the Homestake Pass.\\*
 \qquad  \qquad  \qquad  \qquad  \qquad Necropolis\\*
Hemmed by a penumbra of low foothills, Butte\\*
Dreams of its copper kings, its Mycenaean past, the past\\*
That crumbles with each dawn.

So each dawn proves lethal,\\*
Bleaching the brindled underworld of its deities.\\*
Yet the blue light leaching slowly westward is still\\*
Tentative as it fingers the glyphic landscape:\\*
The trees with bark like scales,\\*
The boulders hunched in the gloom\\*
Like spectral dinosaurs.  The road goes up\\*
And up, ascending in a series\\*
Of mesmerizing curves, threatening\\*
Narcoleptic trance, involuntary return\\*
To that city of shades, catacombed dreams.

When we crest the mountain, we leave\\*
Those shadows behind. The valley before us---\\*
A gleaming basin of honey-colored light---\\*
Will flare in our windows as we descend.\\*
Yet even that light is rooted in darkness\\*
And unfolds its scrolling doppelganger\\*
Over our heads by evening. We're forced\\*
To seek shelter, \& I drift off\\*
To sleep on the rocky motel bed,\\*
Conjuring chimera, the contours\\*
Of moments rekindled out of the blur.

As if they were the traces\\*
Of a lover's face, yearned for,\\*
Irrecoverable.
\end{verse}
