\PoemTitle{Lives of the Saints/Lives of the Terrorists}
\label{ch:lives_of_the_saints}
\settowidth{\versewidth}{For their faith, for their love, for a promise of Heaven}
\begin{verse}[\versewidth]
1.\\*
Nothing is a matter of life \& death.\\*
Not this ordinary day, not the grey road swerving\\*
Up to the college, not the dull faces of the 8 o'clocks\\*
Opaque with boredom.   No wonder

I yearn for the world of Giotto's frescoes, Byzantine\\*
Severity scaffolding all that grief.\\*
And joy. Joy \& grief they do not fear\\*
Though each wracks them.

2.\\*
Martyrdom.\\*
There's a matter of life \& death.\\*
Who are they, the martyrs, the men\\*
Who uncover terror in our hearts\\*
Only to go unnamed to Heaven?\\*
God touches them with tongues of fire.\\*
God speaks through them of His fatal\\*
Meaning.\\*
                 But the translation\\*
In their mouths is ash.

3.\\*
Who were the martyrs of my childhood?\\*
The Catholic saints flayed \& broken\\*
For their faith, for their love, for a promise of Heaven\\*
Fiercer than their pain.\\*
The ferocity of their deaths---\\*
Racked, pierced, crucified---

Not their fear---\\*
They are drained of it, filled\\*
With a terror that amounts to joy.\\*
But their tormentors \& executioners, still\\*
Ordinary, still earthbound,\\*
Still skirting awe, \& afraid\\*
Of the gulf of agony that opens\\*
The saints to Heaven.

4.\\*
They say that very young men make the best\\*
Assassins, terrorists, martyrs, their worlds\\*
Absolute, too pure\\*
For life.\\*
If Heaven is that realm\\*
Where no flower can wither, as Herbert\\*
Claimed, then this earth is the terrorist's\\*
Hell, the realm where no absolute can resist\\*
Decay: to die young\\*
May be to live\\*
Inviolate---\\*
The one\\*
Way.

5.\\*
Stone.\\*
That's my heaven.\\*
Somewhere on the stony crest\\*
Of a moor arching its spine into the slate\\*
Sky, I could lie down, imagine\\*
The world seared to its elements:\\*
Sky, wind, rain, stone. Each\\*
A tidal\\*
Force re-

Calling me always to a world\\*
Not  human, not human-scaled, without\\*
A human face or meaning.\\*
So I resolve to live\\*
In a stone house, a house\\*
Risen from the stone under-\\*
Lying the scrabbled slope \& thickset\\*
Heather: I will become\\*
Stone myself, scoured, scraped\\*
Down to what will endure. But\\*
Nothing\\*
Endures. Everything\\*
In time erodes, corrodes, suffers\\*
Change. Endure:\\*
The root is dur, hard.\\*
Must I be hard?\\*
Will I endure?\\*
What corrodes me?\\*
What wears me down?  Grief.\\*
Love. \\*
Heaven.\\*
All this stone.

6.\\*
Giotto's saints \& mortals\\*
Look hewn. Not impassive\\*
Like stone---but rather\\*
Monumental, stilled by the gravity\\*
Of their lives, of events weighted\\*
By our grief, by our long\\*
Study \& meditation, by\\*
Our longing, our many\\*
Deaths.

It's not that Giotto is perfect.\\*
It's not that we must study\\*
His works.\\*
                Giotto is not perfect.\\*
His paintings are not perfect---there's great\\*
Awkwardness in the poses, the faces.\\*
They're full of effort \& pain.\\*
Or effort \& strain, yearning in the face\\*
Of terrible brutality, senseless death,\\*
Senseless life. Or the appearance\\*
Of senselessness. Stars\\*
Prick open in the firmament, one by one,\\*
Here as anywhere. Constellations\\*
Arrange themselves overhead\\*
And here, as everywhere, saints \& mortals\\*
Raise their faces into Heaven's\\*
Crystalline spheres in wonder\\*
At the star's wheeling\\*
Impersonality, at the music\\*
They almost hear.
\end{verse}
