\PoemTitle{Naturally Death}
\label{ch:naturally_death}
\settowidth{\versewidth}{Ribs like ruined cathedrals rising in the deep summer fields.}
\begin{verse}[\versewidth]
It's spring \& so naturally death\\*
Is everywhere. The drive to work---\\*
An hour each way, morning \& night---\\*
Is littered with bodies. The red-eyed\\*
Surprised `possums, of course, \& deer\\*
Still graceful, like drawings of themselves\\*
Asleep or leaping or glancing back\\*
At where they've been, not where\\*
They meant to go. By summer's end I'll see their bones\\*
Ribs like ruined cathedrals rising in the deep summer fields.

But the muck of death catches me at every turn \\*
This spring: over the smeared raccoon, I observe, the fur\\*
Waves serenely as the uncut grass on graves. Skunks\\*
Fill cubic acres with their stench \& march out\\*
To die in droves, it seems, on back roads \& highways.\\*
The carrion crow, stubbornly tugging some ruby knot\\*
Of brain or intestine, is nailed \&\\*
Splayed in blood \& sand, one wing lifting\\*
In the car's drag. And this morning

On one curve, laid quietly in the gravel,\\*
A small cat, eyes closed: clotted \\*
Blood stains the lavender collar.\\*
Its motionless form is blue with shadow\\*
Except where the first rays catch the frost:\\*
Spikes of light, coldly\\*
Beautiful.  Beyond---\\*
Down the hill sweeping toward town---\\*
The last squares of window-light fade\\*
Into the dawn.

\vinphantom{Into the dawn.} The houses go dark\\*
And ghostly figures emerge, husks of shadow.

Above, the old orchard on the ridge appears\\*
A line of Cyrillic glyphs against the gold

Sky. I am carried away from that\\*
Inscription by the drive to work

Even as the cross-vaulting of night\\*
Dissolves.   The harsh wash of day-

Light bleaches the world\\*
Into flat sameness.

But when night returns in the west\\*
It is edged with the same gold,

Streaks of blood---\\*
And I drive toward it.
\end{verse}
