\PoemTitle{Physics Lesson}
\label{ch:physics_lesson}
\settowidth{\versewidth}{Sense or logic, at least of any kind that might apply}
\begin{verse}[\versewidth]
I show my son how the world is organized---\\*
When it is organized, or when\\*
Or where the scaffolding shows\\*
Through. The lilac leaf is easy.  First\\*
The tiny fisted bud, reddish purple.  The leaves\\*
Each day a little more clearly leaves.\\*
Until this morning, after a heavy night\\*
Of rain edged with thunder,

Isolate leaves stretch loose, are\\*
Singularly green. Each reaches singly\\*
Into the light, although after this\\*
Morning's keen examination we may never\\*
See each so keenly again: they'll blur\\*
Into the swell of greenery that swallows all\\*
Detail by mid-spring.  The sharp air\\*
Of late March will soften,

And the barbiturate scent of nodding\\*
Lilac blossoms, so obviously \\*
Lovely and phallic, will ask us only\\*
To inhale and sleep.  It is luxuriously\\*
Easy to miss the quadrate flowerets\\*
That build the whole.  And yet it may be\\*
Self-indulgent as well as difficult\\*
To puzzle the dual\\*
 \qquad  \qquad  \qquad Nature\\*
 \qquad  \qquad  \qquad  \qquad of the undivided

Light that falls\\*
 \qquad  \qquad over the whole:\\*
Photon \textit{and} wave:\\*
 \qquad  \qquad     the laws of physics\\*
Are unrepentant.\\*
 \qquad  \qquad     They refuse to make sense.\\*
Yet I persist in explanations of the surface\\*
Minutiae to my son, and try to sleep, and cannot,\\*
So cut a slew of lilac for the living\\*
Room, the bedroom, the dead\\*
Hot attic.  Perhaps one day I'll waken

Quickened with comprehension:\\*
The laws of physics are the lines\\*
Of Escher's cities, his serene houses,\\*
Towers, porticoes---perfectly rational\\*
In their details, impossible\\*
Overall. To my Magellan, these\\*
Are outposts beyond common\\*
Sense or logic, at least of any kind that might apply

To the cosmos set out by Euclid.\\*
But our backyard is Euclidean, in its outlines\\*
If not its essence, and the stars can be said\\*
To be arranged as figures---heroes and maidens,\\*
Queens and monsters---and those figures signify\\*
Narratives, meanings, that revolve and\\*
Rise and set in a space and time\\*
Supra-human in scale, though reeled in

Season after season.  But in the small\\*
Solar system centered on my son\\*
Dawn reveals a planetarium\\*
Crashed in the garden, spheres and planetoids\\*
Now soccer balls, tennis balls, beach\\*
And basketballs, the occasional orphaned \\*
Marble, glinting like a sprite's eye\\*
Under the lilac bush.
\end{verse}
