\PoemTitle{Reading Before Sleep}
\label{ch:reading_before_sleep}
\settowidth{\versewidth}{Of ordinary days, ordinary thoughts.  There death}
\begin{verse}[\versewidth]
At nightfall, after dinner, we read nature\\*
Books, my son and I, and in them death\\*
Turns up with surprising frequency.\\*
Tigers eat gazelles. Tornadoes eat skies,\\*
Eat houses.  The skeletons of diatoms\\*
Fall in silent rain on the ocean floor,\\*
And earthquakes, over millennia,\\*
Digest continents.     We don't worry.\\*
It's all so far away, another place\\*
Or time. \\*
\hspace{2\vgap} But snakes. Snakes\\*
We think we know:\\*
Emerald or coral, king\\*
Or king cobra---\\*
Hood flaring like Pharaoh's---\\*
Black and venomous, the black\\*
Mamba, or mildly\\*
Striped and harmless,\\*
The grassy garter snake,\\*
Disappearing, always disappearing

Just as we approach.  Tapestried\\*
Or dusty, damascene or scaly tweed,\\*
Twisting, plaited, all, lethal\\*
Or milk-blooded: snakes\\*
Almost seem familiar.\\*
But snakes in dreams\\*
Don't even seem familiar,\\*
Are not natural, \&, noosing \&\\*
Un-noosing in their loose twists,\\*
Become ancient glyphs,

\hspace{4\vgap} Kanji, \\*
Ink-strokes; a draft or tracery, perhaps\\*
Of the rummage \& debris of the unconscious\\*
Beyond grammar, without law, yielding\\*
From the undertow of sleep, pure image, \\*
Pure form:\\*
\hspace{3\vgap} In dreams\\*
Snakes are a glamoured species, an essence\\*
Like a scythe, gleaming among the stems\\*
Of ordinary days, ordinary thoughts.  There death\\*
In hallucinatory guise appears---\& again\\*
Appears: a scribble of lifeless trees, a lost

Presence, flames on the loom, knot \\*
In the blood: snakes\\*
In dreams are a species not to be found\\*
In natural histories, in bedtime reading\\*
Illustrated \& explained, even the gory \\*
Parts explained.   In photographs\\*
They do not photograph\\*
And their bite\\*
Ends\\*
Explanation

Of every kind\\*
Ends\\*
Speech.
\end{verse}
