\PoemTitle{Segue, Elision}
\label{ch:segue_elision}
\settowidth{\versewidth}{\vinphantom{Slow motion.} When we stop, it is evening. There's a hotel}
\begin{verse}[\versewidth]
Corroded by the eons, the infernal\\*
Cairns \& anvils of the badlands dwindle\\*
Into more ordinary landscape: flat\\*
But uneven, as if hammered

Then abandoned.  The empty prairie,\\*
Riven by the interstate, slopes\\*
Imperceptibly toward the horizon:\\*
A ridge of light like bone. A red truck, tropical

Shadows under the eaves of noon, relieve\\*
The monotony of these generic\\*
Cumulus, this austere grassland muffling\\*
What dead, what bones fluting their loss?

The clouds drift \& ravel, but in fact we move\\*
Too swiftly to perceive that slow\\*
Slow motion.\\*
 \vinphantom{Slow motion.} When we stop, it is evening. There's a hotel\\*
Just where we expected one to be.

Harold \& the purple crayon come to mind.\\*
Isn't it wonderful how desire unfolds\\*
The world we desire before us?\\*
I must have wished for this room

Orderly, motionless, intimate---yet\\*
Cleansed of its ghosts, a chamber haunted\\*
Only by the chambermaid, who leaves her traces---\\*
A glass in wax paper, cool sheets, virgin

Soaps in their cerements set on black\\*
Marble in a moment she already\\*
Doesn't remember?  This morning she smoothed\\*
The depression left by the dark

Body dreaming its dark dreams all\\*
Night in this bed. The dark body I follow,\\*
The dark body we each dream \& fail\\*
To know, waking in the dark, to remember.
\end{verse}
