\PoemTitle{\textit{Thatcheria Mirabilis}}
\label{ch:thatcheria_mirabilis}
\settowidth{\versewidth}{Can't, won't, eat. \qquad And the gastropod,}
\begin{verse}[\versewidth]
The shell, Thatcheria,\\*
Seems unnatural,

Architectural,\\*
Sculptural, a

Spiral\\*
 \qquad Staircase

 \qquad Sharply keeled,\\*
Built to code

 \qquad But not\\*
Nature's code.

 \qquad  \qquad Yet\\*
This must be

Precisely nature's code, even\\*
To the shell's architecture, a kind

Of visual play on the\\*
Double-helix, albeit twisted shut

At the one end, wedged\\*
Open at the other. \qquad Like

The spider's web or the \\*
Swallow's nest, the shell

Is\\*
Constructed

As the web\\*
Is woven, or the nest scaffolded:

Without thought,\\*
 \qquad  \qquad A drive

To make, without comprehension\\*
Of making: no debate in the bird, no

Wonder, no pause to admire\\*
The work.  The swallow cuts

Half-spirals through the clouds\\*
Of insects I can't see, certainly

Can't, won't, eat. \qquad And the gastropod,\\*
Mirabilis, cements its incremental

Whorled masonry\\*
Too slowly for observation. \qquad It

Does not observe itself, does not miss\\*
The self it lacks, does not

Compare \& contrast. \qquad Comparatively\\*
Stronger than steel cable, more delicate

Than the stiff grass it drafts\\*
Into service as a loom: the orb-

Weaver's silk is tacked,\\*
Looped, suspended, spiraled

Methodically: perhaps accidentally\\*
Strung with dew, teardrops 

Of light at dawn.  The cranefly\\*
Mummified, sucked dry, knows

Nothing of this wonder: the wonder,\\*
Mirabilis, is all 

In my mind.
\end{verse}
