\PoemTitle{The Cold: Four Paintings and a Failed Taxonomy}
\label{ch:the_cold}
\settowidth{\versewidth}{For the one glittering scrap the cold might yield.}
\begin{verse}[\versewidth]
1. Bruegel: The Return of the Hunters\\*
~\\*
I know this landscape, thousands\\*
Of miles and hundreds of years\\*
Distant.  Beneath a wintry overcast\\*
The pale grey-green of weathered copper,\\*
The hunters return, hunched\\*
Against the weight of snow and sky.

Their town is unnamed, its brick houses,\\*
Brick inn, brick church now probably long decayed\\*
Or pummeled into dust by one or another\\*
Of Europe's internecine wars.\\*
Peasants, townspeople like Bruegel's\\*
Plowed under like grain.   But here,

In the shelter of this frame,\\*
All turn to their work; none gaze out\\*
Beyond the picture frame, none pause\\*
Or pose, none break from their group.\\*
The winter requires concentration, its demands\\*
Bow their backs: not one stands straight.

Just behind the three hunters, two figures\\*
Bend over an outdoor fire---not a forge\\*
As I first thought: they singe a pig.\\*
A child in an apron watches. In the middle\\*
Distance, a woman carries faggoting \\*
Across a double-arched bridge. Below

There are skaters, lone and coupled,\\*
On the greenish ice, and further still\\*
A man driving a loaded wagon\\*
Toward the alpine crags, heavily white.\\*
Above all, crows, opportunistic, scan\\*
For the one glittering scrap the cold might yield.

2. Bronzino Translates Machiavelli

Around the same time, in Italy\\*
Agnolo Bronzino painted the icy courtiers\\*
Of Cosimo de'Medici's vip'rous\\*
Court.  These women and men appraise\\*
Us, dismiss us, dismiss each other.  Insects.\\*
The uninflected glance

Sums us up.   Their faces\\*
Chill as enamel, pale against\\*
The velvets and satin, hair\\*
Strangely polished, metallic.\\*
Their hearts are talonned.\\*
Their backs prickle with fear.

They horrify themselves\\*
And, so pinned, look to us\\*
Frozen.

3. Portrait of Bartolomeo Panciatichi\\*
                by Bronzino

Now this, this is Hell,\\*
Quite human after all.\\*
It does not burn.\\*
There is no fire.\\*
There are ashes\\*
Where once there was flame.

4. Bosch: Garden of Earthly Delights

Bosch's Hell is cold, despite the flames.\\*
Of it, or in it, nothing human---\\*
Not these, these elfin\\*
Men \& women, foolish, awed,\\*
But without passion, unmarked\\*
By the animal heat of evil\\*
Or good.

Now all they know is suffering.\\*
Even that is mute. There are no screams.\\*
Their eyes are wide, but theirs is the empty\\*
Eyed gaze, the stiff-limbed pose of puppets\\*
Without heat or hate, the anguish\\*
Of any passion. Never in them\\*
The heart consumed, not destroyed:\\*
Their Hell is ice.

5. Animal Kingdom

                                  Who comprehends\\*
Hell? Bosch, Dante, Blake,\\*
Cotton Mather among others\\*
Have raised that etched light, revealed

The demons they could imagine:\\*
Porcine, reptilian, bestial. Always less\\*
Than human---or more: Fallen\\*
Angels fallen into form, collapsed

Radiance knotted into flesh, crumpled visages\\*
Of despair and rage. Rage\\*
The prison, infernal, rage the animal\\*
Emotion.  And despair, the lair,

The cavern, the abyss no animal\\*
Inhabits. Only the one rational \\*
Animal who knows the mind\\*
Is its own place.   So Hell is carried

To the earth's four corners, the corner\\*
-Less sky, terrible, in truth\\*
Infinite, the sky only man\\*
-Kind knows. That Hell. That Heaven

Over earth.
\end{verse}
