\PoemTitle{The Age of  Dinosaurs}
\label{ch:the_age_of_dinosaurs}
\settowidth{\versewidth}{Can we know it? \qquad If                                                       }
\begin{verse}[\versewidth]
Was that world so pure?

Was there no tragedy, only ravage \& raw\\*
Hunger?     No pity, no hope

Without language, not even terror\\*
Without thought   and no thought

In the reptilian brain, just\\*
Blind sight, neurons firing in darkness.

Is consciousness like light?   If so,\\*
It is a light that oddly muddies

The world, silts the eye\\*
Even as it promises revelation.

With thought comes comprehension\\*
And terror. Terror, hope, \& pity

Swell the heart, \& with a full heart\\*
What can we know, \& how

Can we know it? \qquad If                                                       \\*
Dinosaurs survive as birds

Plumed, taloned, \& their reptilian\\*
Brains persist in us, plumed,

Taloned, unconscious, unconscious\\*
-Ly directing our dreams, then whose

Dreams are we dreaming?\\*
What are we \& where

Do we begin \& end?   The viper\\*
-Fish---needle-teeth clustered

In its gaping skull-grin---should be\\*
The very image of death. It is not.

That is life, the urge to live, urgent\\*
Even in the inky cold depths.

Does this creature ever rise\\*
Toward light? \qquad Milky eyes

Bulge forward---what do they look on?\\*
What do they see that light

Never sinks to,   \& how\\*
Do they see it?   That knowing---

Does it become knowledge?\\*
Does it lie, curled \& knotted,

Coiled in our brainstems, snaking\\*
Through events to become

The story?
\end{verse}
