\PoemTitle{Triassic Sandstone}
\label{ch:triassic_sandstone}
\settowidth{\versewidth}{                                  So read the tracks made by this balletic}
\begin{verse}[\versewidth]
    There, preserved\\*
In the red sandstone of the Connecticut Valley,\\*
Are the tracks of an animal that forgot it ran 

Even as it ran---or almost as soon---but now we\\*
Watch it, try to imagine it running---tens \\*
Of millions of years later.

Everything it looked on has gone.\\*
And many many millions of mornings have dawned\\*
And passed, many many millions of days

Stretched \& opened \& then slipped themselves under\\*
The billowing folds of dusk, of midnight, of 3 a.m.,\\*
Before these times had names, or time

Was thought of.     Millions\\*
Of millions\\*
Of nights, blue-black, without a heaven to curtain them,

Became, in effect, all the same night. Became \\*
Night. Stars shifted, perhaps, but for eons, nights \& stars\\*
Left uncounted, for eons stars not mapped

In constellations, for eons\\*
No eons to speak of, no memory, no speech.\\*
                                  So read the tracks made by this balletic

---Coelophysis, was it?---\\*
 In the tens seconds it dashed\\*
Across red sand. What one in a million

Criss-cross of accidents---

Weight \& stride of the dinosaur,\\*
Weather, wind, volcanic eruption, drought---\\*
Left undissolved this brief, quick message

Now telegraphed effortlessly, as it seems, across\\*
Millennia?  How odd to imagine, reconstruct, see\\*
The eyes of a creature that can't return our gaze, nor ever look

On a landscape other than this one, spiked green\\*
With scale trees \& seed ferns \& club mosses; scouring\\*
Rushes, towering forty feet, queerly out of proportion

To the animals below.  Foetal \& stumpy\\*
Or stitched together out of the wrong parts, pig\\*
-Eyed \& fierce: over them in the skies without heavens

The days \& nights interleave in rapid succession\\*
Like a deck of cards, expertly shuffled.\\*
Hundreds of millions of days   of nights  

Each forgets itself\\*
Even as it balances on the edge\\*
Of the next, the rim of dark or light.
\end{verse}
