\PoemTitle{A Shadowed Poem}
\label{ch:a_shadowed_poem}
\settowidth{\versewidth}{Your body, like a piece of furniture, like the bed itself}
\begin{verse}[\versewidth]
\hspace{0.45\versewidth}I

Your body, like a piece of furniture, like the bed itself\\*
Is still. As our suburban lawn at summer dusk\\*
Is still: ragged with grass and shadow.

Your shoes lie empty on the cold linoleum.\\*
Curled on my bed, barefoot,\\*
In recurring dreams you return

To streets brilliantly patterned\\*
As frescoes, painted so as to preserve the illusion\\*
Of presence and depth, monumental architecture

Diminishing toward the imagined horizon.\\*
Waiting for you to catch up,\\*
Your mother stands laughing

With a neighbor.  Sweat\\*
Diamonds her brow: like\\*
One of Masaccio's holy figures, mortal

And hot-blooded, she casts a shadow where she walks.\\*
Sun-blind in the dark entrance-way:\\*
The white-washed walls

Look blue.  You lean against the cool plaster\\*
Watching her.  She unpacks the groceries, sets\\*
Tulips in a glass.  Her speech

Still lilts with the broken cadences\\*
Of a foreign accent, the last echoes\\*
Of a foreign childhood.  She smiles at you.

She smiles at you. She smiles at you. You awaken\\*
Asking ``why\\*
Do her hands not singe

The tulips? Why does she think me\\*
Still\\*
A child?

\hspace{0.45\versewidth}II

One evening in March---the soles of your feet\\*
Burning from shopping all day---you fall\\*
Exhausted on my bed and curl\\*
Into sleep deftly as a child.

 \qquad  \qquad  \qquad The moon

Walks on long adolescent legs\\*
Through your dreams\\*
Through the streets\\*
Of the corroded city.

 \qquad  \qquad  \qquad Once

Cancer gnawed at your breast\\*
As night\\*
Gouges the moon.  The \\*
White-jacketed surgeon

Healed you with his knife\\*
Closed the falcate incision\\*
With catgut.  For a month in the hospital\\*
Your head swam in morphine and you

Could not speak.  Dad and I stood\\*
In silence by your bed\\*
While your tongue \\*
Lolled like a puppet

On slack strings.\\*
On the third day after surgery\\*
The surgeon unwrapped the swaddling\\*
That bound your side. \qquad Examining

The red stitches, the crimped flesh, he seemed\\*
Confused: was it\\*
The cancer or your life-

Blood, life force\\*
That grew turgid \qquad like a sheaf\\*
Of monstrous

Tulips thrust\\*
In the narrow glass of your body?

\hspace{0.45\versewidth} III

 \qquad  \qquad Laced\\*
Into bed by tubes, restraints, pain

Your mother strains against gravity as against\\*
The undertow of her opiated

Blood.\\*
Her feet

Are unmarred by surgeon or cancer, marble-\\*
Smooth, bridal \qquad as if prepared

For a journey.\\*
You wash them \qquad silently

Comparing them to the flayed\\*
Feet in your anatomy text. \qquad  You

Imagine.\\*
You imagine

That your breath enters her---\\*
Surgical---

Healing---\\*
You imagine her

Walking barefoot on the summer lawn\\*
A girl in the northern provinces. \qquad Wedged

Between brick apartment blocks near Boston\\*
New grass shines


With a chemical heat.  Enveloped\\*
In a photosynthetic fire

Unconsumed\\*
The grass

Is a caul on her face---\\*
Early spring

1948.

\hspace{0.45\versewidth} IV

 \qquad  \qquad Swept to your bed\\*
By a flood of bad dreams   I awakened

As you lifted me from the delta\\*
Between you and dad \qquad  \qquad where

Tossed like wild grass seed\\*
By the unbridled river

I'd settled at midnight.\\*
Alluvium deposited on the far shore

Of my bed twenty\\*
Years later slowly

Gathers into a body\\*
---Yours--- \qquad clay

Features dense with sleep.\\*
Dreams:

Ambush:

Childbed:\\*
 \qquad ``The place

For lying in wait\ldots .''\\*
Lochia

Washed your crescent thighs\\*
Six times \qquad now

The roots of new grass\\*
Flicker in your dreams like the legs

Of your adolescent daughters\ldots .

\hspace{0.45\versewidth} V

On the wall I've pinned an anatomy chart: \\*
Stained glass harlequin\\*
It has entered your dreams

It walks the eroding \\*
Streets   like a saint, mortal\\*
And defiant. \qquad At its feet is written:

``Numbered, numbered, weighed, divided.''\\*
Your body, enveloped in membranous\\*
Fire

Unconsumed\\*
Rises through the river of dreams\\*
To surface on my bed late

One March evening in Worcester.\\*
In the window a scimitar \\*
Moon dangles above your head---

Teeters on the uneasy boundary\\*
Of your dreams: toy\\*
Sword, paper cutlass

Rippling into the fire\\*
That takes a thousand forms:\\*
Feral gymnast

Falciform bone\\*
Wing\\*
Fetus
\end{verse}
