\PoemTitle{Alas Poor Hamlet}
\label{ch:alas_poor_hamlet}
\settowidth{\versewidth}{\vinphantom{In philosophy} Well, not that philosophy.\quad Mine}
\begin{verse}[\versewidth]
The days draw in. Summer birds begin to rise\\*
In migratory flocks. The redwing disappear\\*
Mid-feud, even as their territorial cries\\*
Remain behind: How long did I hear

Echoing arguments long past? Absences pre-occupy me.\\*
As good a way as any to escape death, I think:\\*
Think of something else. Or pretend to meet\\*
Death half-way: get all Hamletty, pause, drift, link

Dark weighty thoughts together, smile appalled\\*
At what Hamlet can't un-see, though fleeting:\\*
The skull beneath the skin. Quite chap-fallen:\\*
Alas, poor Yorick's grin is death first greeting

And last word: Hamlet yields up soul \& story\\*
To a stage awash in blood---yet in the end, glory

Dimmed not by death, but doubt: thinking\\*
\textit{Too precisely} parses the whole into gears \& parts,\\*
Splinters all to atoms, protons, quarks, shrinking\\*
Action's scope to Planck's scale. The shards,

Re-collected, never again cohere.\\*
Nor does Horatio's antique Roman virtue\\*
Offer comfort: to stoically persevere,\\*
Glance neither left nor right, subdue

Longing, anguish, regret--bury even joy\\*
In philosophy---\\*
\vinphantom{In philosophy} Well, not that philosophy. \quad Mine\\*
Is ragged as the jay's: in murderous duet

With mate she cracks the dawn, shatters every toy\\*
Of quiet or false calm---and, when she dines,\\*
Dashes chickadee from dewy bough \& shreds the suet.
\end{verse}
