\PoemTitle{Archaeopteryx}
\label{ch:archaeopteryx}
\settowidth{\versewidth}{does not answer my questions. \qquad  \qquad But}
\begin{verse}[\versewidth]
1. \\*
She imagines a Cretan summer\\*
In New England.  Above her in the raw\\*
Morning sky the clouds kaleidoscope:\\*
Continents surrounded

By seas glittering with the scales\\*
Of mermaids with green\\*
Eyes and tails. \qquad  \qquad She tail-\\*
Spins from the porch, leaps

To the precipices of air, scattering\\*
Clouds of brown grasshoppers\\*
Across the scorched lawn. \qquad Tropical

And lush, the colors of her feathers flush\\*
Beyond the black outlines crayoned\\*
In scalloped rows. \qquad Chameleon

 \qquad In scaly wings she\\*
 \qquad \quad Wheels and soars\\*
 \qquad \qquad Glamorous as\\*
 \qquad \qquad \quad Icarus 

Before his plunge into the gaping Aegean.

2.\\*
She snakes out\\*
Into the humid summer night.\\*
The window frames her parents and baby sister\\*
Gathered in the living room, bathed in the light

Of the black and white TV.\\*
The darkness is pinioned to her\\*
Like a cape, a wing, a secret\\*
Identity. \qquad I

Am her moon-fringed shadow trailing\\*
Through labyrinthine bushes late\\*
One summer night. \qquad She \\*
Does not answer my questions. \qquad  \qquad But

Hushes me.  Later\\*
She tents the bedclothes over her head\\*
To protect her from the looming\\*
Presences, the draped

Shadows of the bedroom.  Their\\*
Hair hangs down in silver\\*
Streams; their eyes and teeth\\*
Glitter like coins or moonlight

In her Daedelean dreams.
\end{verse}
