\PoemTitle{As a Child}
\label{ch:as_a_child}
\settowidth{\versewidth}{Was the lull in the storm: the wind-whipped clouds}
\begin{verse}[\versewidth]
1.\\*
When I was a child, I was a bird.\\*
My father's blue eyes were the sky where I

Sailed, loving him\\*
As one may love a natural force: impaled

Even in the high \\*
High heavens on caution \& 

Awe. His storms abandoned me, featherless--\\*
Too human for flight--

Amid the splinters of our house. Cleaved\\*
Of a sudden in half, exposed

To the wind, the torrents of slivered\\*
Rain. Held spell

-Bound by the eye of the hurricane,\\*
I named it:

Love. \\*
Love

Was the lull in the storm: the wind-whipped clouds\\*
The sobbing sky

At bay. My father's eyes\\*
Were sky-blue: when the skies

Blackened, I spilled from the heavens\\*
Like a bird, my

Heart snapped\\*
Shut in mid-

Flight.

2.\\*
Her eyes were green oceans.\\*
My mother knew stories, secrets, love, delight.\\*
She was the sun we orbited,\\*
The moon that silvered our dark dreams.

Walking away, I walked\\*
Toward her. Cupped\\*
In her lap, measuring my small, water-\\*
Smooth hand against the sand

-Stone of her palm, I wished for her heart,\\*
Her secrets. I did not ask:\\*
Why does the harbor of this love yield\\*
Us to storm?

3.\\*
In purple-shaded jungles\\*
My sisters, my brother, \& I\\*
Linked fingers \& day-dreamed:\\*
Fat green snakes looped about

Gaudy blossoms, dragonflies\\*
Darted like needles through the glinting\\*
Air. But our comicbook panorama gave way\\*
To my father's blustering

Narrative: The moon\\*
Was a ghostly galleon, tossed\\*
Upon stormy seas. Even as he read,\\*
His wrath rose, his voice

Filled with a howling wind, encircling\\*
The house. The splintered moon\\*
Vainly gathered \& regathered\\*
The fragments of its perfect face

From heaving waters: night\\*
After night\\*
We gazed, rocked\\*
By waves even as we lay

Under cover, abed.\\*
Until that dawn when I raised\\*
White sails to catch the \\*
Wind that sped to me

From the green islands south\\*
Of our imagining. And one\\*
By one\\*
Called you after me, my siren's voice

Emptying the house,\\*
Leaving our wrathful father--\\*
Our mother, the sandstone figure\\*
Still at the center of his whirlwind rage.
\end{verse}
