\PoemTitle{Azrael}
\label{ch:azrael}
\settowidth{\versewidth}{So I will become the runner\ldots I will become the runner}
\begin{verse}[\versewidth]
\hspace{0.45\versewidth} I

Late July:\quad summer has passed\\*
Its zenith:\quad the sickle descends

Towards autumn.\quad But while it's still summer\\*
The athletic fields behind the college

---Unmown, sea-green---\\*
Are empty by 

Dusk. \\*
The horizon

Darkens last of all.\\*
Tonight---for a long while tonight---

The horizon wriggles on the surface of the river\\*
Like a fire-green snake suspended

Between origin and destination.

\hspace{0.45\versewidth} II

A hundred yards away from me\\*
Across the river\\*
A figure in white:

A man in jogging gear:\\*
Baggy white pants and tee-shirt\\*
Blue shoes with yellow lightning

Stitched on.\\*
Poised in the middle of the chalked \\*
Arena, he stands as if

Waiting.  Does he see me?  Does he want me\\*
To leave?  He begins\\*
His warm-up exercises:

Each movement\\*
Petals out of the previous movement\\*
Counterpoints it

He stretches\\*
Toward the sky, toward the grass,\\*
Away from me, toward me.

At this distance, in this light,\\*
His features are a lunar blur.\\*
He could be a woman.

\hspace{0.45\versewidth} III

He moves\\*
Slowly 

Gazing\\*
Straight

Ahead\\*
Seeming

Not\\*
To move

For\quad long\\*
Seconds

Then

Dipping fluidly\\*
Right or left.

Hands clasped above his head\\*
(Do they seize a sword?)\\*
He pauses\qquad then

Gazing sightlessly across the green river\\*
Extends his arms\quad opens\\*
Into a new

Pose:\\*
Palms upturned\\*
Arms outstretched

Like the crossbar of the scales\\*
That weigh the souls of the dead.

\hspace{0.45\versewidth} IV

So I will become the runner\ldots I will become the runner

I will become the man in the field at dusk\\*
But I will remain a woman

Dressed in baggy white clothing,\\*
Utilitarian clothing, meant for running

And I will wear those famous blue running shoes\\*
Yellow lightning, wings of fire, stitched to my heels

I will become the man in the field\\*
The runner at dusk, a woman

Poised on the edge of summer\\*
Looking across the river

Where the further bank is obscured\\*
By the smoky air; where a woman

Or a man stands, watching me---\\*
She appears

A motionless pillar\\*
And I am the woman or man she watches

As I grasp the sword above my head, stretching\\*
Or as I scythe the grass with my fingers

Dreaming\quad at the culmination of each\\*
Movement:\quad pausing on the inhalation

Then flowing out with my breath\\*
Across the field, across the field

I am the runner.
\end{verse}
