\PoemTitle{Blood Trail}
\label{ch:blood_trail}
\settowidth{\versewidth}{Invisible and cold and silent.   I'll remain}
\begin{verse}[\versewidth]
It's not that I didn't love you.\\*
Damn you!

It's very cold where I am---\\*
Bleak---\\*
The stars are out\\*
The trees stripped\\*
Bare, black\\*
Like bars---this\\*
Is my solitary\\*
Confinement, out in the open\\*
Where no one dreams\\*
Of finding me.

\ldots .This hillside's steep.\\*
No one overhears\\*
These words to you\\*
These poems \ldots .

I wish you could hear them.\\*
But I can't\\*
Come near you.  This night\\*
Holds me back.   The black cold, the steely\\*
December stars, Orion the warrior\\*
Bar me from you.\\*
Yes, I think he lost his humanity\\*
Up there, whirling through the eons.\\*
He's forgetful.

I haven't forgotten.

But it's cold where I am,\\*
Very cold, and solitude\\*
Is a drug that makes the cold\\*
Beautiful, desirable\\*
Like a weapon, a blade\\*
That penetrates at once and takes\\*
The heart out whole---

Next to that, what can you offer me?

Love? It weights my heart\\*
With a future \\*
And a past\\*
And the dead---\\*
My dead, the men I carry\\*
Everywhere---\\*
Crowd me\\*
Till I cradle them, one at a time\\*
While the fear drains from their eyes\\*
Until finally all light and warmth\\*
Decays and they're cold,\\*
They return\\*
To the realm of shadows\\*
Invisible, but present, here, now,\\*
Everywhere\\*
In the submerged \\*
Landscape of night, the dim \\*
Hills and streets, shimmering\\*
Winter fields and the thin houses\\*
Built to shelter us from dark and cold.\\*
And yes they follow me, sniffing out\\*
My blood trail even into daylight\\*
So that one, at least one,\\*
Always follows at my heels\\*
Reeking of death\\*
And love.   It seems easier\\*
To meet them here\\*
Where they remain\\*
Invisible and cold and silent.   I'll remain\\*
Silent with them\\*
While early morning\\*
Darkens around us\\*
And the watery stars\\*
Fade, out here\\*
On the last \\*
Promontory of night          where \\*
No one \\*
Dreams.
\end{verse}
