\begingroup
\PoemTitle{By the Laughless Rock}
\label{ch:by_the_laughless_rock}
\settowidth{\versewidth}{Resurgence of sap in their winter veins.\quad Her mother}
\begin{verse}[\versewidth]
\hspace{3\vgap} In spring, she steps

Out of tunneled Hades, returns\\*
To her mother, the natal embrace\\*
Of earth and sky.\quad The grey saplings shiver,\\*
Anticipating the annual\\*
Resurgence of sap in their winter veins.\quad Her mother\\*
Releases the cool ferns from their tightly\\*
Curled dreams; newts wriggle free of the loose mud\\*
At pond's bottom:\quad tiny clouds of silt\\*
Billow in their wake.  In spring, her lithe

Heart is a salamander, slender and\\*
Amphibious, living on the border\\*
Between terror and flight.\quad In spring, she slips\\*
From the scaly grip of her chthonic life\\*
Emerges glistening and nimble, listening---\\*
Listening for the voice of her\\*
Nascent lover.\quad The maple saplings\\*
Bow to the gathering wind,    then lift their arms:\\*
Like a child's scrawl, the curlicues

Of opening leaves feather the supple\\*
Branches.\quad Persephone steps\\*
Into the living\\*
Field, the filaments of grass\\*
Electrified in the commingling\\*
Of chlorophyll and light.\quad Persephone\\*
Returns to the meadows\\*
Where she was born, where once she played: no\\*
Longer a girl, her eyes are dark as caves.
\end{verse}

\renewcommand{\poemtoc}{subsection}
\clearpage

\PoemTitle{Halifax, 16 October 1918}
\label{ch:halifax}
\settowidth{\versewidth}{Between contractions.\quad France was darkening}
\begin{verse}[\versewidth]
My mother's mother stepped out\\*
Of her street clothes, lay\\*
On the high and narrow bed, counting\\*
The red leaves falling past her window

As the midwife counted the minutes\\*
Between contractions.\quad France was darkening\\*
On the other side of the earth;\\*
Turning away from her: the blossoming pear

Orchards, her maiden self. \textit{Poirier,}\\*
Her maiden name, \textit{pear tree}, anglicized\\*
To \textit{Powers}. Now her Irish husband curses\\*
The gravity of pears in autumn, swaying

Earthward over grass matted gold.\\*
She cries out she cries\\*
Out \textit{Maman!}\quad  Framed in the white\\*
Hospital window, the slender maples

Shiver; crowned in red, rattle\\*
Grey knuckles against spangled ribs.\quad Release\\*
The last leaves in a vivid flush.\\*
Born against the scarlet tide: \textit{Dark Rosaleen}:

The black-haired daughter, green-eyed,\\*
Upon the sea-foam of her mother's\\*
Thighs plunged into the new\\*
World as it opened, wide
\end{verse}

\clearpage
\PoemTitle{Providence, April 23, 1952}
\label{ch:providence}
\settowidth{\versewidth}{Bed, these torn cries, this red-faced squalling}
\begin{verse}[\versewidth]
White clouds somersaulting in the blue\\*
Sheer sky; below, the lucent tongues\\*
Of new grass whisper against the\\*
Soles of my mother's shoes.\quad She walks

Across the hospital lawn, up\\*
The marble stairs overhung\\*
With forsythia.\quad Inside, the pure\\*
Volumes and white edges of her

Hospital room ruptured\\*
By a spray of forsythia, dis-\\*
Located in a pear-shaped vase.\\*
Beyond, the network of black-topped

Streets all leading to this\\*
Bed, these torn cries, this red-faced squalling\\*
Daughter in the crescent\\*
Of light pooling between her thighs.

On the third morning, my mother stepped\\*
Into her street clothes, whispered\\*
\textit{Alice Frances---Mother---\\*
This is our daughter.  The waxen leaves}

\textit{Of her small lungs, her leaf-\\*
Green eyes, the garland of her new\\*
Life is your name writ\\*
In green flame.}
\end{verse}
\endgroup
