\PoemTitle{Desire}
\label{ch:desire}
\settowidth{\versewidth}{The searing, distant god?      And if you}
\begin{verse}[\versewidth]
Your outspread hand, outstretched\\*
Eclipsed the side of my face, swept\\*
The cloud of red hair from my cheek.\\*
Whiskey and cigarettes: I relished\\*
The taste of that first kiss,\\*
Gilded my lips and tongue with bourbon\\*
Licked transparent gold\\*
Wherever you gave way.\qquad Later,\\*
Naked, I rested beneath your crazy\\*
Grandmother's parti-colored Afghan\\*
Netted where I lay--on the edge\\*
Of your never-made bed, smoking\\*
Cigarette after cigarette you sat\\*
Hunched, absent, stroking my exposed\\*
Throat, looking and refusing to look\\*
Into my exposed face, you struggled\\*
Not to say what the late hour,\\*
Rolling Rock and whiskey\\*
Split from you. Ragged\\*
Split psyche, black mirror of the\\*
Black oval, the new moon etched,\\*
Almost invisible, in the bruised\\*
Glittering sky. By dawn, dredging\\*
Kisses from each other's lips,\\*
We pitched into crazy-quilt sleep,\\*
You shuddering as you entered\\*
Your dreams. I nestled in your\\*
Iron arms: you embraced but did not\\*
Hold me. Gold light embroidered my\\*
Lashes in the morning; I floated\\*
Above our sheet-sculpted bodies, not\\*
Desiring answers or knowledge, but\\*
Desiring\\*
That this be an image of death:\\*
The brazen morning light, the March gales\\*
Battering at the blinds--\\*
You and I in our crystals,\\*
Barely breathing but entwined.\\
Asleep--I hovered above the bed,\\*
Not dreaming--in sleep your face\\*
Looks like the child Apollo's\\*
--Was he ever a child? Or eternally\\*
The searing, distant god?\quad And if you\\*
We're Apollo-- or any other\\*
God--that would explain your hair,\\*
A platinum corona--but\\*
Your eyes so black and sweat-stung--why?\\*
With eyes so dark and wet, why\\*
Did you shudder as you entered me?
\end{verse}
