\PoemTitle{The Dogwood Tree}
\label{ch:dogwood_tree}
\settowidth{\versewidth}{Sit down on the river-bank, watching}
\begin{verse}[\versewidth]
\hspace{0.45\versewidth} I\\*
~\\*
A figure\\*
In the distance\qquad across\\*
The still\\*
Fallow spring

Fields.\qquad Too

Far to tell\\*
Whether approaching\\*
Or departing or\\*
Standing

Still.

The sunlight\\*
Is pale, the wind\\*
Tosses my dress, my red\\*
Scarf\\*
In scrolls.\qquad I walk,\\*
Stuttering your empty\\*
Name.\\*
For

A moment

It fills, radiant,\\*
With the\\*
Light\\*
Rippling\\*
On the river bottom, then

\hspace{0.45\versewidth} II\\*
~\\*
\ldots then? Nothing.     Red\\*
Berries\\*
Drop from a thorn\\*
Bush; I\\*
Sit down on the river-bank, watching\\*
Nothing, the sun\\*
Goes behind a cloud.  My\\*
Birth-month is approaching, April\\*
Is my birth-month, Shakespeare,\\*
The tradition\\*
Says, was\\*
Born on my\\*
Birthday.\qquad And Prokofiev\\*
too, and Planck, father\\*
Of the very small\\*
And very\\*
Constant\qquad born\\*
On my birthday. \qquad In\\*
April, late\\*
April, the\\*
Forsythia\\*
Blossom, the\\*
Magnolia blossom, tulips\\*
\& lilac \\*
Overtake the \\*
Roadside, sweeten\\*
The grey\\*
Suburban\\*
Yards.\qquad My heart\\*
Flutters, empty, billowed\\*
Open\\*
Like a sheet\\*
Hung on the line, caught\\*
Up\\*
In the spring\\*
Breeze, the virginal\\*
Light.\qquad \qquad Your name

Is an echo, a\\*
Whisper\\*
In cellars, doorways, white-\\*
Scrubbed\\*
Hallways, the vestibule\\*
Of my ear.\quad Here

Is the light, the\\*
Distant figure\\*
Still\\*
Paralyzed \\*
At some\\*
Indeterminate\\*
Distance.\quad Perhaps\\*
It is not human: is it\\*
The play of \\*
Light\\*
On a tree\\*
Stump, a tree?\quad Its crossed\\*
Branches\\*
Waving\\*
Like arms?\quad Towards

Me?

\hspace{0.45\versewidth} III\\*
~\\*
Light. Light\\*
Impossibly nailed to a tree.  Light\\*
~\\*
Seeking the obscure\\*
Face of\\*
Love, as\\*
Abstract\\*
As God's.\quad Light\\*
Spilling blood from the\\*
Air, the bushes, the\\*
Fields\\*
Newly \\*
Plowed, un-\\*
Planted.\quad This

Is the light's\\*
Quick\\*
Penetration\\*
To the\\*
Depths where\\*
White\\*
Roots\\*
Weave \& weave.\qquad I

Pluck the green lush\\*
Foliage\\*
From a face

It remains \\*
Concealed\\*
In the tangled\\*
Jungle-\\*
Green, green\\*
Vines\\*
Blossoms\\*
Roots

Piercing


The resistant\\*
Loam.\quad I\\*
Scrape at the earth with\\*
My nails, impatient,\\*
Tired.\quad A little grit,\\*
Rubble, comes away \&\\*
I think of the\\*
Childhood story: God, \\*
Molding\\*
Adam \& Eve\\*
From blood-soaked\\*
Dust, whittling\\*
Time \& space\\*
From never, nowhere, never,\\*
Never.\\*
Yanking the first\\*
Startled \\*
Pair of us\\*
From our earth-black\\*
Amnion\quad into\\*
The precipitous\\*
Days \\*
Of his\\*
Love, so like\\*
Wrath, into \\*
The blinding\\*
Light\\*
Of

Day.\\*
Over

My hands\\*
Ripples\\*
The same\\*
Light.\quad By


This same\\*
Light,\\*
I examine the rubble: the\\*
Foliage tenaciously\\*
Clinging\\*
To that hidden\\*
Desirable\\*
Face, I\\*
Impatiently\\*
Push aside, up

Root\qquad to reveal

What?\qquad This net of\\*
Pulpy\\*
Gleaming\\*
Roots?  Naked,\\*
Twitching toward the earth\\*
From which they\\*
Were torn?

The magnolia 

\hspace{0.45\versewidth} IV\\*
~\\*
Tree\\*
Is bowed\\*
With the weight of its\\*
Immaculate\\*
Blossoms\\*
Earthward.\hspace{3\vgap} The dogwood

Has not yet flowered.
\end{verse}
