\PoemTitle{Dream House}
\label{ch:dream_house}
\settowidth{\versewidth}{Statuary, lunar flowers, tined, silvery, scentless.}
\begin{verse}[\versewidth]
In dreams, houses\\*
Always have one more room\\*
Than I remember\qquad sometimes\\*
An entire suite\qquad a forgotten\\*
Attic or corridor\qquad doors\\*
Opening unexpected where

The wall had been smooth or a mirror\\*
Failed to reflect me.\\*
If there is piano music there is no piano\\*
When I uncover the space enclosing\\*
The fading echoes.  The room itself\\*
Is an echo of some other place

And as I try to map\\*
The Mobius topology that sweeps me\\*
Out of the room in the same motion\\*
I enter, my eyes stray to the spines\\*
Of books sheathed in dust, their gold leaf\\*
Titles indecipherable, their authors

A people angular as cranes.   The piano\\*
Is hushed but from that hush arises\\*
Other sibilance.   Voices under the rain.\\*
Thorn bushes rasping on windowpanes\\*
Moonlight scraping over seashells.\\*
If there are windows they reveal

The room looks on sky; if there is landscape,\\*
It is alien, a garden of fossil bones, broken\\*
Statuary, lunar flowers, tined, silvery, scentless.\\*
It will be night when I arrive,\\*
Night when the room jells\\*
Out of moonlight, its contents and contours


Coming into focus minute by minute.  Every\\*
Dream minute could prove\\*
A thousand years on waking\\*
Or a thousand years, passed through\\*
In a single night of dreaming, forgotten\\*
At dawn.  But day after day, night

After night, I forget one world\\*
Upon waking into the other. The gears\\*
Of sleep and dreams do not mesh\\*
With the tinsel machinery of day.  And no\\*
One dream prepares me for the rooms\\*
I discover in the next. Dream

After dream, I do not recognize\\*
This wing of the house, this garret, these\\*
Stairs ascending to a blue-floored\\*
Refuge.  Yes, I think the floors\\*
Are always blue, as if\\*
Tiled with sky, but the walls are unstable

Ever on the verge of dissolution\qquad like\\*
Fish\quad schooling\quad dispersing\quad in iridescent\\*
Flash and shimmer---in that\\*
Dissolving always there\qquad always\\*
Just elsewhere dream room\\*
Cartesian coordinates crumble

The architecture of memory proves\\*
Kaleidescopic\qquad and the mosaic\\*
Of stories and gaps\qquad a frieze of images\\*
At odds with expectation\qquad is strangely\\*
Unfamiliar, though I have for the moment\\*
Given them symmetry, and said they were mine.
\end{verse}
