\PoemTitle{In Time}
\label{ch:in_time}
\settowidth{\versewidth}{The Christmas wreathes and ribbons: they flutter}
\begin{verse}[\versewidth]
1.\\*
No one takes down \\*
The Christmas wreaths and ribbons: they hang\\*
Riding the cold wind in the desolate park\\*
As winter drags, then brightens\\*
Into spring.   But no\\*
One yet strolls the cement walks, least\\*
Of all\\*
The man

Who was or is or appears\\*
To be my lover. Shall I wait\\*
For him to return, with his child---\\*
Or to rise, with the new year\\*
 A new man---\\*
Long hair wet with rain, skin\\*
Rinsed with light,\\*
Bearing

The garland, the crown\\*
Of flowers, the bough twined\\*
With serpent, heavy with fruit?

2.\\*
The passing sacrificial season\\*
Bestows promises\\*
Opening yielding\\*
Further promises: endless\\*
Deferral of the desired\\*
Paradise where no flower\\*
Can wither: Herbert's eternal\\*
Life is death.  But here\\*
In every season\\*
The greenery interweaves the sharp \\*
Palings, the fleshed\\*
Body veils the pale\\*
Ribs, sharp teeth---\\*
The mortal threat \\*
Of the mortal body.\\*
\vinphantom{Of the mortal body.} Here\\*
In this sublunary \\*
World crucified\\*
On the twin axes \\*
Of time and place the blossoms\\*
Must wither\\*
Before the bough bears\\*
Fruit.

3.\\*
No one takes down \\*
The Christmas wreathes and ribbons: they flutter\\*
In the summer breeze as the park crowds \\*
With children.\qquad They are weighted\\*
If only slightly, by the gravity\\*
Of their guardians: parents and sitters\\*
Clump in their midst\\*
Insisting on a center to the chaos\\*
Spiraling out from them. My lover\\*
No longer my lover\\*
Is not among them\\*
Nor will he ever again\\*
Appear in quite this \\*
Light.

4.\\*
Now\\*
His arms are empty\\*
And I do not wait for his\\*
Return.  In the end\\*
He proves a man\\*
Entirely\\*
And only.

5.\\*
The desolate park\\*
Is dusted with a light snow\\*
That suggests blossoms.\\*
New lights are spiraled\\*
Round the central tree to celebrate\\*
The solstice, birth\\*
In the dark of the year.  There is\\*
A time for mourning but it is\\*
Not as spring approaches, wings\\*
Of light on\\*
Her heels.
\end{verse}
