\PoemTitle{In the Season without Mercy}
\label{ch:in_the_season_without_mercy}
\settowidth{\versewidth}{ \qquad  \qquad There's the danger, too, of remembering}
\begin{verse}[\versewidth]
1.\\*
All spring and summer birds wove nests\\*
In the ivy scrolling the southern face\\*
Of my new apartment.  I followed\\*
The arc of their dark bodies into the blue\\*
And their swooping return; the cries of chicks\\*
And the squabbling of mates alike\\*
Were music to me.  Song and warmth and late\\*
Long dusks: summer's golden net\\*
Entangled but could not shelter us.\\*
Now that it's November and the bright leaves\\*
Fallen and only the pink stems left to shelter\\*
The sparrows and finches from the early\\*
Failure of light and the bladed cold, I turn\\*
Inward, contemplating shelter,\\*
Shelter and solitude and loss.  There's not \\*
Much solitude: my daughter makes quick hash\\*
Of that, demanding company, food, a song.\\*
So we sing.  And when our harmonies \\*
Have faded, and she's asleep\\*
The apartment grows still, except sometimes\\*
For a bird rattling in the eaves, or neighbors\\*
Trundling down the stairs, whistling.\\*
I work. The computer clicks and hums\\*
Neither content nor discontent---\\*
But I wander from my text, turning\\*
Over and over the fragments of my marriage---the vows\\*
Broken before I broke them, the intimacy all\\*
Scripted in therapy sessions.  His face averted, \\*
He claimed the rights and role of husband\\*
With a vengeance.  I owed him\\*
Everything I offered, and nothing\\*
Moved him to delight, a kiss, surrender---\\*
That brief shelter offered without thought.

2. \\*
 \qquad  \qquad  \qquad  \qquad  \qquad Where\\*
Do the birds shelter on a night\\*
Like this one? Fierce cold rain, a vengeful god\\*
Somewhere overhead, rolling thunder, flashing\\*
Blue-white fire about our heads?\\*
I turn off the computer and check\\*
My daughter, then darken all the rooms\\*
And watch lightning play across the veiled face\\*
Of infinity.\\*
 \qquad  \qquad There's the danger, too, of remembering\\*
Only what he remembers, only the last \\*
Years, and not the first dance. We did love\\*
Once: I trace in my daughter's features\\*
Mother and father.  She imagines herself\\*
A bird---chickadee or wren---and curls\\*
Against my breast, strokes my hair\\*
Inquisitively, declares it soft, takes flight\\*
For other realms, but always returns\\*
Not doubting love, though she's begun to fear\\*
Its loss.  What do I tell her?  Should I\\*
Lie, promise her shelter from death and loss,\\*
When every day renews the loss\\*
Of her faraway father?  He calls,\\*
He calls and visits, but some mornings\\*
Her limbs are heavy with the loss of him,\\*
Inexplicable, though she tries to wrest\\*
Meaning from the fragments: she recites\\*
Stories that slide into myth, draws\\*
Maps of his fairytale life, and her journey\\*
To recover him.  He's always at the end\\*
Of the road, smiling, in a house as tall\\*
As he is.

3.\\*
 \qquad  \qquad Sometimes\\*
These last few years seem the tale\\*
Of people who lived a century\\*
Past---and I've studied them, dead,\\*
Awkwardly posed in sepia and ivory\\*
Unreadable anguish shadowing their eyes.\\*
Frozen for a split second, they pause for eternity\\*
At the beginning and end of time\\*
Where neither future nor past\\*
Offers shelter\\*
And no birds sing.
\end{verse}
