\PoemTitle{Litanies}
\label{ch:litanies}
\settowidth{\versewidth}{             ---The woman with the black silhouette of crows on her soul,}
\begin{verse}[\versewidth]
#1\\*
Litany for the Woman with the Rose

The woman with the rose\\*
The peach-colored rose,\\*
The wrong color for winter.

The woman with the manic eye\\*
The depressive vision\\*
The dead blood.

The woman who watches as the silence is gathered up,\\*
Who watches as the lights go out, one by one,\\*
Absolute as the stopped eye.

The woman who grabs the dark road by the tail\\*
And of course cannot let go.\\*
The woman with the dead white rose.

      ******

The woman who holds the rose\\*
The orange-scented rose\\*
The summer smell unsettling in the winter streets.

The woman silently gripping her world\\*
As it drains of reality and grows flat:\\*
The level glare of glassy eyes and black windows.

The woman who draws the flat black line\\*
On the electro-encephalogram,\\*
Who says nothing, monstrous or common\\*
Who answers with absolute silence.

     *****

 \qquad *****

The woman who draws the rose\\*
The gut-soft, gut-pink rose,\\*
The petals wilting and bleaching in the cold dusk.

The woman who clearly delineates everything in her world\\*
In the black noose of a line\\*
So that she may know it---

Whose eye is alternately enticed and stripped raw\\*
By the knife-edge of the line,\\*
The line which links her to life\\*
And separates her from it.

Her vision is raw but ecstatic.\\*
She is the woman who draws the charcoal rose on brown paper.

#2\\*
Litany for the Woman with the Orchid

The woman with the very perfect orchid\\*
The orchid made of words, words glassy and bald as eggs\\*
The orchid marble-smooth and moon-like in its distance\\*
From her heart, her grasp, her green-gold eyes.

The woman with the smile set in the face\\*
The face once thought beautiful (in her youth).

The woman whose dark eyes seek an exit\\*
In marble flowers, lunar coolness, blue distances;\\*
In the long road away, tomorrow, black\\*
Tomorrow.

The woman who watches the stars spiral out from her\\*
Who watches the light swim about her (without touching\\*
Her---)\\*
             ---The woman with the black silhouette of crows on her soul,\\*
The woman who talks to these shadows.

#3\\*
Litany for the Woman and the Ocean

The woman who seeks to wound herself with the ocean\\*
(The ocean which bleeds dark visions onto the shore\\*
The ocean which wounds itself and heals itself, which is constantly\\*
Singing, roaring, diving into itself.)

The woman who seeks to obliterate herself with the ocean\\*
The woman whose mouths\\*
Open as sea anemones on the ocean floor\\*
Whose ribs arch into rocks, barnacles, shells\\*
Whose words dissolve into crackling whistling whispers static.

The woman who is rinsed with salt down to the bone,\\*
The woman who melts to music.

The woman who dissolving with each breath, finally knows\\*
That the sky hemorrhages into ocean\\*
That blue shades into sea-green\\*
That wave bleeds into wave, rolling toward the shoreline\\*
Which wracked by these, flexes and shivers\\*
Like a mirage.

#4\\*
Litany for the Woman in the Museum

The woman with the poppy\\*
That red queen, the executioner:\\*
The poppy whose mouths have burst\\*
Bloody but laughing.

The woman saucer-eyed, dark-eyed, arrogant and beautiful.\\*
The woman who seems to see and cannot\\*
The woman who once saw.

The woman whose face is painted on her funeral shawl,\\*
Who gazes at me across two thousand years---\\*
The woman who knows me though she may deny me.

The woman haunted by Anubis,\\*
The woman whose heart is weighed in the scales---

The woman inscribed from head to foot\\*
In curses and blessings,\\*
Rituals and songs,\\*
The chants of the dead.

The woman at whose breast gnaw\\*
Jackals and falcons, gods and lovers\\*
Song upon song, the cries of the living.\\*
The woman who asks me to let her die.

#5\\*
Litany for the Woman in Spring

The woman with the apple\\*
The apple red hard cold\\*
Fruit of her lips her hips her thighs:\\*
The apple glistening \& sighing\\*
And changing color.

The woman with the apple in her lap\\*
Apple of earth \& blue rains, harder than bone\\*
Redder than fever or death.

The woman with the lap full of fruit\\*
The belly of melon, milk, heapfuls of wheat.

The woman whose cedar-scented skin\\*
Tingles and aches at the teasing of spring blood---\\*
Whose nipples, under her white linen shirt\\*
Harden like apples\\*
When spring breezes touch them.

The woman who watches as\\*
Red buds swell on the trees\\*
Slender limbs begin to ache with green\\*
And the bark of the apple tree rouges delicately\\*
Amid the golden grasses and blue March skies.

The woman who watches her own body\\*
Swell, ache, tingle and rouge\\*
With hard sweet fruit with love\\*
The rock red blossom the pink-petaled moon-fruit---

The woman who feels at last\\*
The surging ease---\\*
Moon \& apples gasping in her throat\\*
White release, red knowledge:\\*
Life flickering, murmuring\\*
In the tightening womb.
\end{verse}
