\PoemTitle{Lung Cancer, Stage IV}
\label{ch:lung_cancer}
\settowidth{\versewidth}{Works at giving himself laryngitis       everything}
\begin{verse}[\versewidth]
Is not the best thing to have ever\\*
Happened to me     but I am\\*
Chastened by the thought that everything\\*
Has been.          Now\\*
As I write these lines\\*
And the rooster a cornfield away\\*
Works at giving himself laryngitis       everything\\*
Has been.      Everything is locked\\*
In the past.    When

I next read these lines       likewise\\*
The past will be just\\*
As inaccessible, an attic\\*
Stuffed and crammed with everything\\*
Imaginable---and unimaginable---but without\\*
Door or window of any kind.\\*
You can stand for eternity\\*
On its doorless porch      knocking knocking\\*
On the blank façade---but like a black

Hole that's got it all wrong, even as the past\\*
Balloons, its gravity will only cast you off, not\\*
Draw you, spaghettified, over\\*
The invisible eventless event horizon\\*
Into its singularity.\\*
So.      I could be dead\\*
In a year. Nobody's saying.\\*
There goes the dauntless rooster\\*
Again, as if he hasn't just

Crowed a dozen dozen times in some\\*
Quixotic affaire d'honneur with the sun\\*
That the sun knows nothing of.\\*
Who threw down their gauntlet\\*
First? The sun blinks at the question\\*
And rises. Time\\*
Now\\*
To go for a walk\\*
Perhaps to join Stevens'

\\*
Sailors and tigers in their\\*
Red weather, perhaps\\*
Merely to see\\*
What rustles the corn, who\\*
Pops out of the cornfield\\*
At moonfall, and whether\\*
Their eyes reflect purgatories\\*
Or ordinary\\*
Light.
\end{verse}
