\begingroup
\PoemTitle{Marilyn Poems}
\label{ch:marilyn_poems}
\renewcommand{\poemtoc}{subsection}
\PoemTitle{Hourglass}
\label{ch:hourglass}
\settowidth{\versewidth}{For the blue and white bus.\quad The city}
\begin{verse}[\versewidth]
A stream of photons\\*
Curbed by gravity\\*
You take form:\qquad Egyptian

Lotus eyes, oval\\*
Face, a body that tells\\*
Time.\qquad You died

When I was ten:  now I\\*
Am a woman and like you\\*
I can't sleep.\quad Shrouded

In rain, we hold\\*
Hands as we wait, stranded\\*
For the blue and white bus.\quad The city

Is inscribed: neon, scribbled\\*
Trees, the runic alphabet\\*
Of weathered brick,

A newspaper\\*
Blown\\*
Against your ankles.

Before we go, tell me:\\*
What is this city, this\\*
Hieroglyphic face?  

I'm a woman,\\*
Like you, I'm not\\*
mortal.
\end{verse}

\clearpage

\PoemTitle{Downtown, Waiting for Marilyn}
\label{ch:downtown_waiting_for_marilyn}
\settowidth{\versewidth}{Shaped into a Cupid's bow.\quad \textit{Marilyn}}
\begin{verse}[\versewidth]
Death could be this red truck\\*
Barreling down on me in the rain.\\*
Marilyn appears at my side, sheathed\\*
In light, her mouth a slash

Shaped into a Cupid's bow.\quad \textit{Marilyn}\\*
I say \qquad\qquad \textit{You're late.}\\*
She murmurs---apologies?---and this\\*
Is mysteriously seductive---

Not her tigress-\\*
Like symmetry, not\\*
Her royal gaze summing

Up the truck, the wet\\*
Asphalt, the random\\*
Bits of color and movement---

\textit{I am not sure what she has said:}

She compels me to imagine---\\*
Imagine!\quad In the second\\*
Split before the truck hits.
\end{verse}

\clearpage
\PoemTitle{Photograph}
\label{ch:photograph}
\settowidth{\versewidth}{And desired.\quad Each eyelash is curled and raised}
\begin{verse}[\versewidth]
Bloodless and immortal, there you lie:\\*
To a hair---as you are imagined\\*
And desired.\quad Each eyelash is curled and raised

Over your iron-blue eyes---like the\\*
Oars of Cleopatra's golden barge,\\*
Raised and lowered in unison.

Flooded with light, from every angle\\*
Marilyn you take form on the light-\\*
Sensitive film, in the photographer's

Eye?\quad Atropos spun\\*
In a mesh of light:  if\\*
Not visible,

I am nought.
\end{verse}
\endgroup