\PoemTitle{Phoenix}
\label{ch:phoenix}
\settowidth{\versewidth}{                                         There---}
\begin{verse}[\versewidth]
1.\\*
The bus is still cold\\*
When it reaches our stop\\*
And dark          the aqueous streets

Empty      the sun\\*
Not yet up.           Rising\\*
We file into the same seats

Every morning           every morning\\*
The bus carries us east\\*
To work.    The familiar route

Unreels     scene\\*
By scene     in our windows\\*
Framed: elementary

School       playing fields     corn\\*
Graveyard      clover     skeletal\\*
Farm buildings submerged

In shadow.

2.\\*
                     Steel\\*
Arc              concrete

Pinion       reigned\\*
By tension           bent

Between stone and sky\\*
The bridge

                        Over\\*
Curling black water

Looms.                     Blue\\*
For miles around us          for miles

Around us         boreal\\*
Landscape       pillars

Of Lombardy poplar           sunk\\*
In mountainous               shadow.

                               Grinding\\*
Into low gear         the bus

Accelerates        uphill             pauses\\*
On the crest---

3.\\*
Tapestry of flame, the sky\\*
Ignites: royal purple, cochineal,\\*
Tin-white, gold---the light

Clears our sheathed faces       paralyzes\\*
Time.         The two men\\*
Seated behind me

Stammer into silence.\\*
                                         There---\\*
On the mountaintop

Coruscating plumage\\*
Impossible---

4.\\*
Impossible to describe\\*
Afterwards.\\*
                             The bus descends

A ball of fire\\*
Balances on the horizon, breaks\\*
Free---rises, yellow

Disk on the blue sky.            By noon\\*
The two astonished men\\*
Will be unsure of what they've seen---

By evening, each of us will wish\\*
For a handful of tail-feathers---\\*
Souvenir, talisman,

Fire---

To carry home to lovers\\*
And children, or the empty rooms where\\*
Once they were

Or someday they\\*
May be: by such light\\*
We will know them.
\end{verse}
