\PoemTitle{Sheer Poetry}
\label{ch:sheer_poetry}
\settowidth{\versewidth}{                 Unnecessary to say what they meant---}
\begin{verse}[\versewidth]
1.\\*
I was your husband---you owed me\\*
Faith and patience and love\\*
But you drove me out---I fled\\*
As from a burning house---burning! I say---

You seared us both in the fire,\\*
Broke our marriage vows---\\*
And all for that liar! Telling tales\\*
Like the Ancient Mariner on a drunk.

The war. His unspeakable sins, his\\*
Secret wounds. The lost \\*
Daughter of his first marriage---all

Woven with a golden thread, a scarlet\\*
Fringe, a honeyed tongue. Poetry!\\*
What do you know about him,

Really?

2.

You don't\\*
Wonder at my hatred---you leave it\\*
Unexplained to your friends\\*
Who notice and swerve\\*
From my glare.  Oh their lives are tidy and bright\\*
As TV laundry, and sometimes\\*
We too pretend our lives are divided\\*
Into episodes brought to a climax every week\\*
Then folded away.\\*
As if all this pain were artifice, \\*
As if daylight reveals\\*
Everything, as if what happens in plain \\*
Sight were the whole

Truth.

Surely you know better,\\*
Bitch, screwing\\*
My husband in the evening \\*
And disappearing in the glare\\*
Of sunlight when I approached. \\*
Did your fantasies include\\*
My tears?  My children, \\*
Their bafflement, their mute\\*
Questions as we watched those sitcoms\\*
All those evenings?\\*
My yearning and rage, silenced\\*
By his drunken silence, his averted eyes\\*
At 3 a.m.? What's your true love mean, anyway---\\*
All hedged round by lies?

3.

Yours, his, mine, my husband's:\\*
We were hedged round by lies

Like a ring of flame.\\*
We turned our seared faces from the heat

And sought happiness, \\*
Fragile and momentary

As the cold blue sky\\*
Arched over us, we crossed

The red bridge. We entered\\*
The autumnal forest. Our children

Followed, gathered\\*
Yellow leaves which they held

Up to us, faces bright\\*
In the cool dark woods.

That was our tangled refuge\\*
Interwoven with decay

And our daughters' cries\\*
Of discovery or alarm.

4.

You want me to say I never \\*
Loved her---you want me to lie\\*
Even as you scream Liar!\\*
You don't know what love is\\*
You say. That's not love, that's\\*
Screwing, that's all.

I hate lying. So\\*
I tried not to say anything---no\\*
Explanations, no stories. \\*
But as evening closed\\*
In, and the TV poured forth\\*
Its smothering fantasies

And all the arguments I'd narrowly\\*
Escaped all day began\\*
Crackling and flaring,\\*
I'd bow out, claiming the simple\\*
Need for solitude and wine.\\*
I'm working on a poem I'd offer

Not quite lying, resorting\\*
To metaphor as I slipped\\*
Into the blue-grey light.\\*
My lover was waiting, \\*
Her child quieted, her husband\\*
Barricaded in his new life.

And it was true---she was the poem---\\*
Or we were, together\\*
Murmuring and crying out\\*
And in the long silences\\*
Drawing the silence out\\*
In suffocating kisses, desiring\\*
To drown rather than return

\\*
To the lies we narrowly\\*
Avoided giving voice to---bad enough\\*
To live them. We tried\\*
A hundred ways to smother\\*
Doubt, but most nights\\*
Fell back to kisses broken by tales, the telling

Episodes of our lives told\\*
In a murmur like rosaries or ghost\\*
Stories. \\*
                 Unnecessary to say what they meant---\\*
Enough to recite them in the other's\\*
Presence, between kisses, hedged round

By lies.

5.

Really\\*
How do I know \\*
Anything? Everyone\\*
Tells stories, sings\\*
Sad songs---\\*
                        And maybe he was\\*
The worst of all, the biggest liar,\\*
The saddest.  But certainly

Hs lies---if they were lies---\\*
Were wonderful, entrancing.\\*
I lay shattered in their wake, shattered\\*
Too all the tragic myths underpinning\\*
Memory and desire, shattered\\*
All the statuary from the lost world\\*
Of my marriage. Everything was lost.

And the world was new.  His kisses,\\*
His inexplicable absences, his lies\\*
Forced me to enter the world \\*
Alone, as if clothed in flame,\\*
As if listening\\*
For the voice of an angel I was doomed\\*
Never again to hear.    The lies

---If they were lies---\\*
Were magnificent, Byzantine\\*
 In their severity and splendor.\\*
And they have left me here\\*
Alone and full\\*
Of wonder where the new grass\\*
Grows amid the ruins.
\end{verse}
