\PoemTitle{Sunset on Market Street}
\label{ch:sunset_on_market_street}
\settowidth{\versewidth}{Inside its depths shimmer stacks of bowls like white blossoms}
\begin{verse}[\versewidth]

The sidewalks have grilled all day under a white-\\*
Hot sky: now rain steams from asphalt and rooftop,\\*
The horizon is clearing.\qquad \qquad Sharon

Drags the dog-yellow couch out of the shop\\*
Onto the sidewalk where it's cooler, there's\\*
A breeze.\qquad Across the street, the proprietor of the market

Unlocks his car door: hot air belches\\*
From the black interior: the rain-\\*
Washed sky and tropic sun ripple mirage-

Like above the dusty parking lot.\\*
\textit{Beautiful evening!} the old man declares---\\*
Startled, pleased, Sharon looks up: \textit{Yes} she says

\textit{Lovely},  thinking that the rose light on her leaf-green shirt\\*
Is lovely. \textit{We'll have another just like it\\*
Tomorrow!} he adds

Gesturing with a grin toward the banked\\*
Storm clouds, scudding rapidly south.\\*
Her reading interrupted\qquad (besides

It's growing dark)\quad Sharon fingers a scrap\\*
Of velvet, midnight-blue:\quad fascinated\\*
Drapes it the length of her bare thigh:

\textit{An evening dress\\*
Slit so high!} She shivers, looks around---\\*
Is anyone else on the street?\quad \textit{A dress the color}

\textit{Of midnight!  and between my thighs black curls,\\*
Cochineal lips}\ldots . She grins,\\*
Lifts the coils of dark hair from her neck


Repressing the urge to laugh aloud, alone\\*
Where she sits on the dog-eared couch.\\*
The sun is setting: the junk-shop window

Blazes for a moment: goes dark.\\*
Inside its depths shimmer stacks of bowls like white blossoms
\end{verse}
