\PoemTitle{The Body of the Dream}
\label{ch:the_body_of_the_dream}
\settowidth{\versewidth}{     \ldots . I know that my flesh is made of chemicals and molecules. I know that what I'm seeing is like a dream. I know that consciousness is something totally different but yet in some powerful way connected to this physical/chemical organization\ldots .}
\begin{verse}[\versewidth]
                                        I

     \ldots I know that my flesh is made of chemicals and molecules\ldots .

The sunlight\\*
Catching the\\*
Treetops\\*
In green fire-green

Is a physical\\*
Phenomenon.     Also

Physical:     the water\\*
Petaling out of itself, over\\*
The dam's\\*
Edge.      Physical:      the bright\\*
Wall of air, the pond's\\*
Glassy\\*
Surface,    the mirage\\*
Of shadowy\\*
Fish, spark-\\*
Ling\\*
Beneath the surface.

      *****

Beneath the lightless eye of the electron microscope:\\*
Stippled shadows, the shifting\\*
Configurations of our atoms.   We orbit\\*
A gleaming nucleus of dreams, whose smallest\\*
Component may not be dissected

From the body of the dream.

      *****

Physical:\\*
Phenomenal:

Hands and water-smooth thighs       hands\\*
And the eel of the tongue, slithering\\*
From its dark recess to lick\\*
At the body's\\*
Light.     Physical:

Hands slivering the phenomenal\\*
Light, hands\\*
Parting the water\\*
So that the body may follow.

Physical: the flow of the body\\*
Into itself, the spill\\*
Of light beyond the contours of the flesh.

Phenomenal:   physical:   the dream, the body\\*
Knitting together their fingers    of light \&\\*
Flesh, imagining each other.

                         

                                                   II

     \ldots .I know that my flesh is made of chemicals and molecules. I know that what I'm seeing is like a dream\ldots .

Beneath the mirror-\\*
Smooth surface of the dream,   sleep-\\*
Less fish gleam, darting

Up\\*
From the river's still\\*
Depths,   into\\*
The clear shallows,   the light-\\*
Filled pools.          The river\\*
Is luminous

Yet

Is contained\\*
By the riverbank.   The\\*
Shimmering\\*
Boundaries\\*
Unfold from the river, following\\*
Its motion, created

In the river's shifting edge.

     *****\\*
The momentary construction from a moment's\\*
Vision:   the bright wall of my flesh, here, beneath my hand,\\*
Now\\*
Billowing into petals of light.

      *****\\*
The molecules of oxygen\\*
Suffuse my\\*
Lungs, are\\*
Diffused\\*
Through each respiring cell\\*
Of my body.     I am\\*
A lung, a translucent\\*
Tissue\\*
Dilated with\\*
Blood, flushed\\*
With light.

I know\\*
That my flesh is made of\\*
Flesh, is\\*
A veil cast over a\\*
Veil, is,

Is\\*
Not.

     *****\\*
Appearing amid the slivers of light: a\\*
Bird, a black bird, pecking at the flotsam\\*
Eddying in a corner of the concrete\\*
Levee.     Soon\\*
[new line]

It will see me, up here, watching---

Watching---

And will fly\\*
Away.     There:

Gone.  

                        III

     \ldots . I know that my flesh is made of chemicals and molecules. I know that what I'm seeing is like a dream. I know that consciousness is something totally different but yet in some powerful way connected to this physical/chemical organization\ldots .\\*
                                                                                                    Greg Gillespie

In the black well of the\\*
Sky,   stars

Glitter like the scales\\*
Of vast, dark

Fish, gliding\\*
Spellbound,   form-

Ing constellations.   Here\\*
I am, here

I go, here\\*
I come.     Every dream

Is alchemized from flesh into\\*
Flesh.   Flesh itself a dream:

Flickering in the moment, swaying on its stem \&\\*
Dreaming.    A trillion

Cells in my body, a trillion\\*
Stars in the sky.

A trillion stems of grass, blades\\*
In my heart, flame-

Green tongues in my throat, searing\\*
My lips closed, turning words to ash.

                        IV

Physical

Phenomenon:     this tissue

Of flesh, dreaming\\*
On my luminous\\*
Bones.   Physical:

The skin of light, spun\\*
From my own, the skeins\\*
Of grass, weaving emerald\\*
Fire with

Shadow.     Physical

The white stones, nestled\\*
Among roots in the warm\\*
Loam.   On this earth,

I walk.   Through\\*
Air petaling open like\\*
A lover's thighs:   absorbing\\*
Me, giving\\*
Birth to me.   Shimmering\\*
With my afterimages.     Here

Now

I step into the shining\\*
Envelope of my body, the vessel\\*
Spun from clay: silicates \&\\*
Water, thrown on the wheel, raised\\*
Into a cylinder\\*
Whose mouth encircles darkness, opens\\*
To light.   Spinning

On the wheel of my self, now,

With each breath petaling in\\*
To myself, opening

With each tug of my lungs at the air.\\*
Physical:

This momentary\\*
Convergence\\*
Of chemicals \& heat, light \&\\*
Water, flesh \& blood-\\*
Filled dreams.   Physical:

The revolution of electrons about their\\*
Imaginary\\*
Nuclei.     Physical:     the imaginary

Here \& now, physical, the nipples\\*
Lips \& hands of this imaginary body,   the rip-\\*
Tide of light, the bolting fish caught\\*
In these mercurial\\*
Waters.
\end{verse}
