\PoemTitle{What Birds}
\label{ch:what_birds}
\settowidth{\versewidth}{Shriek and stark blue, white, and black}
\begin{verse}[\versewidth]
Are on the feeder?   On\\*
The porch?   Unthreading\\*
Sunflower seeds from\\*
Raveling grass?     I mean

At my house in Candor--\\*
Coo-ah-coo-coo-coo--\\*
Mourning doves mourn\\*
And squeak--the wings do--

They never oil them!\\*
And they bully the other birds--\\*
Snatch their food--they bully\\*
Each other--and the jays

Have nothing on them.\\*
Crests up, the jays land, eye\\*
Seeds scattered in thick arcs--\\*
Crests deflate, they snatch

Two or three, or half\\*
A dozen; discourteous\\*
To no other bird, they\\*
Retreat to railing, bush,

Some more obscure perch--\\*
Leave the falsely peacenik\\*
Doves to harass each other,\\*
Berate the world

Into accepting them\\*
As what they aren't:\\*
Harbingers of peace\\*
Cooing their Zen-

Quaker mish-mash,\\*
Black eyes not reflecting\\*
The world or love for it:\\*
Their mournfulness

Sends me into a rage.\\*
I'd rather have the raw\\*
Shriek and stark blue, white, and black\\*
--The boldness of the jay--

Who guards herself, takes\\*
What she needs, and arcs\\*
Up into the bowering\\*
Birch, her aerie, her

Momentary shelter.   She knows\\*
No shelter lasts.

The chickadees and titmice I\\*
Love for their bold eyes.   Small.\\*
Tough. Determined: they know\\*
What they want and no cat

Nor mere dove prevents them.   Up\\*
Down   weaving   darting: sunflower\\*
Seed or raw peanut\\*
Is taken.     The doves

Are irritated.   The juncos\\*
Hop about the margins, embroidering\\*
Tiny footprints into new snow.\\*
A downy, head ratcheting

With typical woodpecker worry,\\*
Makes a tactical here, now\\*
There approach; a red-bellied, larger\\*
But with equal nervous caution

Brings up the rear.   Their caution,\\*
Though like my own,\\*
Makes me laugh.   Perhaps noticing\\*
The doves absent, a jay

Returns, zig-zagging\\*
From perch to porch, its now\\*
Thinned necklace of seeds; eyes\\*
Their Morse: so grabs

A very full mouth and\\*
Gulletful, then wings off\\*
To the closest birch--\\*
Her shelter

Of the moment.
\end{verse}
